### 2024-02-21 paper/methods.qmd
imputation

## Corpus assembly

The US General Services Administration receives and publishes public comments on proposed rules using the website Regulations.gov.  Commenters may submit comments by online form, email, or physical mail, and electronic formats allow commenters to include a number of attachments.  Attachments are typically detailed comments or to include journal articles, reports, etc., as supporting evidence.

All agency-published documents and public comments associated with a proposed rule are collected together in a docket, e.g., EPA-HQ-OA-2018-0259 for Strengthening Transparency, available at <https://www.regulations.gov/docket/EPA-HQ-OA-2018-0259>.  Regulations.gov provides an API (application programming interface) via http, meaning a simple script can be used to automatically retrieve ("scrape") the full contents of a given docket.  [A1] used the API to retrieve all publicly-available comments — including comment metadata, comment text, and any attachments — throughout 2020 and 2021, completing retrieving in June 2021.

For Strengthening Transparency, Regulations.gov reports 993,464 "comments received" and 22,390 "comments posted to this docket" and publicly available.  *[TODO: email for clarification on relationship between these numbers]*  It's likely that this disparity is due, in part, to separately enumerating some mass comments that were submitted collectively.  For example, comment 11822 (<https://www.regulations.gov/comment/EPA-HQ-OA-2018-0259-11822>) includes 22,786 individual comments collected by Friends of the Earth (likely using an online form) and submitted as large PDFs, one individual comment per page.

After scraping, our corpus included the text of 22,390 comments and attachments for 20,787 distinct comments.  *[try reruning scraper sometime]*  By number of attachments, the median document had no attachments and the maximum number of attachments was 17.

## Sampling

Initial, undirected exploration of the corpus by [A1] indicated that (a) there was a qualitatively clear division between supporters and opponents of ST, (b) commenters offered substantive reasons for their position on ST, but (c) opposing comments were much, much more prevalent than supportive comments.  The first point meant that manual qualitative coding could be used to identify supporting and opposing comments; combined with the second point, this meant that the corpus was suitable for contrasting the arguments given on either side of the ST controversy.  But the third point meant that a simple random sample would include very few supporters.  We therefore needed a way to oversample supporters before manual coding, that is, before classifying them as supporters or opponents.

Further exploration — and a fortuitous coding error — identified a solution.  Using the `spaCy` natural language processing (NLP) suite and its R API `spacyR` *[versions]*, we analyzed the text of each document into individual words ("tokens") and classified each term's part of speech (noun, adjective, etc.).  Specifically, we applied dependency tagging, which enabled us to assemble adjective-noun bigrams (pairs of words), such as "public health" or "regulatory science."  *[filtering large and small docs; comment text and attachments concatenated]*

At this point, [A1] was working on the project by themselves, and had identified the relationship between science and health as the key research question for the project.  [A1] therefore focused on "science or health" bigrams, that is, bigrams where the noun was either "science" or "health."  [A1] identified 405 "science" bigrams and 225 "health"; 215 "science" and 117 "health" bigrams (332 total) occurred in more than one document in the corpus.  15,846 documents contained one or more of the 332 health or health bigrams.

These 16k documents and 332 bigrams can be represented as a matrix, with documents as rows, bigrams as columns, and each cell containing the number of occurrences of that bigram in that document.  In text mining this is called a "document-term matrix."

Principal component analysis (PCA) is a statistical technique often used for dimension reduction or data compression, e.g., compressing the 332 columns of bigram data down into, say, 5 or 10 columns *[cite]*.  The resulting columns are referred to as "principal components," and are sometimes interpreted as latent variables measured by the observed variables.  For example, in intelligence testing, a method like PCA is used to construct general intelligence, a latent variable, out of observed test responses.  [A1] hoped that PCA would find a principal component that could be interpreted, at least roughly, as latent support vs. opposition.  However, [A1] miscoded the script, compressing the documents rather than bigrams.  While this error eludes interpretation — the principal components represent continuously-varying latent documents? — the first principal component did appear to provide the desired oversampling:  many of the documents that loaded onto one end of the first principal component supported ST.

[A1] began to develop a strategy for sampling from across this principal component.  Drawing the most extreme documents from either end resulted in a large number of nearly identical "form letter" comments, most likely collected by advocacy organizations using online forms.  Using phrases that appeared in these comments and simple text matching, [A1] excluded 8,875 from three different forms from the manual coding sample.  After these exclusions, drawing the most extreme 200 documents (including ties) from each end resulted in *[actual numbers, pre-fixing html issue]* 214 from the "bottom" (apparently supportive) and 201 from the "top" (apparently opposed).  Then a simple random sample of 400 comments was drawn from the remaining 6,561 comments (15.8k that contained one or more bigram, less the 8.9k from identified form letters and the 415 from the extremes).  The resulting sample contained 814 comments.^[While manual coding was underway, [A1] prepared draft code for the imputation and analysis steps described below.  This revealed an error in the part-of-speech tagging:  spaCy assumes it is working with plain, pre-cleaned text, but the comment text can contain HTML, especially `<br>` tags.  After fixing this error, the bigram subset was slightly smaller, and some documents used in manual coding were no longer in the bigram subset.  This is why the data used for imputation and analysis is slightly smaller than reported here.]

## Manual coding

At this point, [A1] brought the other collaborators into the project.  Without informing the other authors of the research question, all authors independently coded each comment in the sample as supporting or opposing ST, or too ambiguous to classify.  Krippendorff's alpha was used to assess inter-rater reliability *[https://link.springer.com/article/10.1007/s11135-004-8107-7]*; this statistic ranges from -1 to 1, with 1 indicating perfect agreement across all raters and 0 indicating "no agreement beyond chance" *[https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4974794/]*.  After an initial round of independent coding, Krippendorff's alpha was 0.86, with discordance on 74 / 814 comments.  We prepared a spreadsheet that contained only the discordance comments and each author's codes.  After discussing a few examples, each author separately reviewed the discordant comments, and submitted revised coding sheets.  After this one round of revision, Krippendorff's alpha was 0.94, with discordance on 32 / 814 comments.  Inspection of the remaining discordant comments suggested they were generally subject to reasonable disagreement.  For example, we disagreed about whether the following comment (EPA-HQ-OA-2018-0259-10925) should be classified as opposing ST or is too general and therefore irrelevant to ST:

> We need the Environmental Protection Agency to protect us, and to not loosen restrictions on air and water quality.  Contaminated air and water are terrible for human health.  The EPA should be looking out for people, not fossil fuel companies.  People with respiratory illnesses are especially vulnerable to COVID-19.  As environmental regulations are rolled back, it could be catastrophic for them, as they struggle to breathe already.  I am a registered dietitian, and my work involves helping people to stay as healthy as possible.  This is hard to do on an increasingly  uninhabitable planet.  Our public agencies should be here to protect and work for us.

Having achieved a very high degree of concordance, we moved on to the next stage of analysis.

## Imputation

To take advantage of the scale of the ST corpus, we used machine learning to impute or estimate support or opposition across *[all]* 20,787 comments, using the `tidymodels` framework *[software versions]*.

To assemble the input data, we combined the manually coded comments with the 8.9k form letter comments, treating all of the latter as opposed to ST.  Because ambiguous comments were extremely rare (33/814), we removed them from data, and defined the classification task with a dichotomous "support" or "oppose" variable designated `support`.  Overall, this gave us 9,655 documents, of which 9,499 (98%) opposed ST.  We constructed 3-1 training-testing splits (7,241 and 2,258, respectively) stratified on `support`: 123/7241 = 1.7% supported ST in the training set, and 33/2258 = 1.4% support ST in the testing set.  The training set was further split for 5-fold cross-validation (5792 and 1449 or 5793 and 1448), again stratifying on `support`.

We defined a model workflow that took the raw text of the comment — comment and attachment text concatenated, as with the NLP tagging — as input.  For processing efficiency, text was truncated after the first 3,000 characters, covering approximately the first 500 words, on the assumption that even long comments would express their support or opposition clearly within the first page or so.  Next, the text was tokenized, stopwords removed *[source]*, and tokens lemmatized.  The resulting token features were hashed to $1024 = 2^10$ features, with signed binary values (0, +1, -1).  Zero variance features were removed.  Finally, because `support` was heavily imbalanced, during training supporting comments were upsampled using simple resampling.  Use of the `tidymodels` framework ensured that this preprocessing pipeline could be applied exactly to both training and testing data, as well as the full corpus, without any risk of data contamination.

Model tuning was done using 5-fold cross-validation.  Because of class imbalance, balanced accuracy and specificity were chosen as the primary assessment metrics.^[In `tidymodels`, categorical outcomes are handled as factor variables, and by default the "positive" level of the outcome is the first level/value of the factor.  Since `support` takes the values "oppose" and "support," the first level (and so the positive) is opposing ST.  Specificity, true negative rate, thus assesses the model in terms of its ability to identify support for ST.]

We initially explored the use of a random forest model with the `ranger` package, tuning both number of predictors and minimum splittable node size.  However, fitting these models was extremely slow, and cross-validation estimates were lackluster, especially for specificity (roughly 70%).  A logistic lasso model with the `glmnet` package *[version]*, tuning the lasso penalty $\lambda$, was much faster to fit and had much more promising cross-validation estimates, with balanced accuracy around 96-97% and specificity around 93-94% in the best-perfomring models.

Based on tuning results, we selected a penalty of $\lambda = 0.0385662$ and fit the logistic lasso on the full training data.  This model had a balanced accuracy of 96% and a specificity of 94% on the testing data.  We then used this model to impute dichotomous supporting/opposing codes to the *[full]* corpus of 20.8k comments, without carrying over manual comments.  After imputation, the corpus contained 14,995 opposing comments (72%) and 5,792 supporting comments (28%).  To allow for ambiguous comments, we filtered down to comments where the model was more than 80% confident in its classification (predicted probability of opposition was either greater than 80% or less than 20%).  This filtered subset contained 14,101 opposing comments (74%) and 4,876 supporting comments (26%), for a total of 18,977 (91% of the comments run through imputation).

### 2024-02-21 scripts/07_check_coding.R

@@ -57,17 +57,18 @@ assert_that(identical(
 
 ## IRR using Krippendorff's alpha
 ## Initial IRR (completely independent, no review) was 0.86
+## After one round of revisions, 0.941
 coding_df |> 
     select(starts_with('support')) |> 
     t() |> 
     kripp.alpha()
 
+## After one round, discordance on only 32 comments
 coding_df |>
     select(comment_id, mode,
            starts_with('support'), starts_with('notes'),
            identical) |>
-    filter(!identical) |>
-    view()
+    filter(!identical) |> view()
 
 # TODO: coding against PCA
 

### 2024-02-21 scripts/08_impute.R

@@ -1,4 +1,5 @@
 library(tidyverse)
+library(magrittr)
 
 library(readxl)
 library(here)
@@ -17,15 +18,17 @@ data_dir = here('data')
 
 ## Load data ----
 ## Form letters
-form_letters = read_rds(here(data_dir, '06_form_letters.Rds'))
+form_letters = read_rds(here(data_dir, 
+                             '06_form_letters.Rds'))
 
 ## Manual coding results
-coded_df = read_excel(here(data_dir, '06_docs_2024-01-16_DJH.xlsx')) |> 
-    filter(!is.na(support)) |> 
+coded_df = here(data_dir, '07_coding.Rds') |> 
+    read_rds() |> 
+    mutate(mode = map_chr(mode, str_flatten)) %T>%
+    {print(count(., mode))} |> 
     mutate(support = case_when(
-        support == 'n' ~ 'oppose', 
-        support == 'y' ~ 'support', 
-        support == 'NA' ~ NA_character_,
+        mode == 'n' ~ 'oppose', 
+        mode == 'y' ~ 'support', 
         TRUE ~ NA_character_), 
         support = as_factor(support))
 
@@ -47,24 +50,6 @@ identical(n_distinct(comments), length(comments)) |>
 
 message(glue('{length(comments)} manually coded documents'))
 
-# ## Document-term matrix on full corpus
-# ## NB already tokenized and lemmatized
-# dtm_corpus = open_dataset(here(data_dir, '03_annotations')) |> 
-#     filter(pos %in% c('NOUN', 'PROPN', 'VERB', 'ADJ')) |> 
-#     count(comment_id, lemma)
-# 
-# dtm_sample = dtm_corpus |> 
-#     filter(comment_id %in% c(comments, form_letters)) |> 
-#     left_join(coded_df, by = 'comment_id') |> 
-#     mutate(support = if_else(comment_id %in% form_letters, 
-#                              'oppose', 
-#                              support)) |> 
-#     select(comment_id, support, lemma, n) |> 
-#     collect()
-# 
-# ## sample: 9688 docs
-# n_distinct(dtm_sample$comment_id)
-
 ## Document text
 get_text = function(...) {
     open_dataset(here(data_dir, '03_text.parquet')) |> 
@@ -84,7 +69,7 @@ text_sample = get_text(comment_id %in% c(comments, form_letters)) |>
                              support)) |>
     filter(!is.na(support)) |> 
     mutate(support = as_factor(support)) |> 
[…]

### 2024-02-22 scripts/09_bigram_analysis.R
map bigram analysis over datasets

@@ -1,6 +1,7 @@
 library(tidyverse)
 theme_set(theme_minimal())
 library(ggforce)
+library(patchwork)
 library(here)
 library(arrow)
 
@@ -8,27 +9,28 @@ extract_noun = function(df, in_col = bigram, out_col = noun) {
     mutate(df, {{ out_col }} := str_split_i({{ in_col }}, '_', -1))
 }
 
+## TODO: write out plots throughout
+
 ## Load data ----
 data_dir = here('data')
+source(here('R', 'load_coding.R'))
 
-coding = read_rds(here(data_dir, '07_coding.Rds')) |> 
-    select(-starts_with('support'), 
-           -starts_with('notes')) |> 
-    mutate(support = case_when(
-        mode == 'y' ~ 'support', 
-        mode == 'n' ~ 'oppose', 
-        TRUE ~ NA_character_
-    )) |> 
-    select(-mode)
+coded = list(manual = load_manual_coding(), 
+             filtered = load_imputed_fltd(), 
+             imputed = load_imputed_full())
 
-nrow(coding)
+map(coded, nrow)
 
 ## Number of docs by support
-count_docs = function(coded_df) {
-    count(coded_df, support, name = 'n_docs')
-}
-n_docs = count_docs(coding)
-mutate(n_docs, share = n_docs / sum(n_docs))
+## Here we only want number of docs *w/ bigrams*
+# count_docs = function(coded_df) {
+#     count(coded_df, support, name = 'n_docs')
+# }
+# # n_docs = count_docs(coding)
+# n_docs = map(coded, count_docs)
+# 
+# mapped(n_docs, ~ mutate(.x, share = n_docs / sum(n_docs)))
+
 
 ## Load bigrams ----
 load_bigrams = function(coded_df) {
@@ -36,33 +38,44 @@ load_bigrams = function(coded_df) {
                       '05_bigrams.parquet')) |> 
         inner_join(coded_df, by = 'comment_id') |> 
         filter(support %in% c('support', 'oppose')) |>
-        collect() |> 
+        collect() |>
         extract_noun()
 }
 
-bigrams = load_bigrams(coding)
+bigrams = map(coded, load_bigrams)
 
 bigrams |> 
-    pull(comment_id) |> 
-    n_distinct()
+    mapped(~ {.x |> 
+            pull(comment_id) |> 
+            n_distinct()})
+
+n_docs = map(bigrams, ~ {.x |> 
+        distinct(comment_id, support) |> 
[…]

### 2024-02-24 paper/methods.qmd
map bigram analysis over datasets
format methods doc for BJPS

@@
+---
+format:
+  pdf:
+    mainfont: "Times New Roman"
+    fontsize: "12pt"
+    geometry: 
+      - "margin = 1in"
+    linestretch: 1.5
+---
+
 ## Corpus assembly
@@
 
-To take advantage of the scale of the ST corpus, we used machine learning to impute or estimate support or opposition across *[all]* 20,787 comments, using the `tidymodels` framework *[software versions]*.
+To take advantage of the scale of the ST corpus, we used a machine learning model to impute or estimate support or opposition across *[all]* 20,787 comments, using the `tidymodels` framework *[software versions]*.
 
@@
 This filtered subset contained 14,101 opposing comments (74%) and 4,876 supporting comments (26%), for a total of 18,977 (91% of the comments run through imputation).
+
+## Text analysis
+
+Text analysis was conducted as reported in the main text, with code written in R *[version]* and making extensive use of the `tidyverse` suite *[version]*.
+
+## Data and code availability
+
+The full data collection and analysis code is available at *[github]*.  To facilitate reproduction and reuse, *[separate data repo]*
+
+- raw ST scrape results? 
+- 03 raw text and annotation files
+- 06 manual coding
+- 08 fitted model and imputed dataset

### 2024-02-24 scripts/09_bigram_analysis.R

@@ -168,23 +168,44 @@ p_df |>
     imap(llr_plot) |> 
     wrap_plots(nrow = 1, guides = 'collect')
 
+# opts = options(pillar.sigfig = 10)
+## For manual, medians are identical bc the median bigram appears in 1 opposing doc and no supporting docs
+p_df |> 
+    map(llr) |> 
+    mapped(~ {.x |> 
+            extract_noun() |> 
+            group_by(noun) |> 
+            summarize(llr = median(llr))})
+p_df$manual |> 
+    complete(support, bigram, fill = list(df = 0)) |> 
+    group_by(support) |> 
+    summarize(df = median(df))
+log1kp(1/622) - log1kp(0)
+# options(opts)
+
 llr_by_noun_plot = function(llr_df, title) {
     llr_df |> 
         extract_noun() |> 
         mutate(bigram = fct_reorder(bigram, llr)) |> 
         ggplot(aes(noun, y = llr, color = noun)) +
-        geom_violin() +
-        geom_sina(aes(label = bigram)) +
+        geom_violin(scale = 'width') +
+        geom_sina(aes(label = bigram), scale = 'width') +
+        # stat_summary(fun = median,
+        #              geom = 'crossbar', 
+        #              color = 'black') +
         geom_hline(yintercept = 0) +
-        scale_color_brewer(palette = 'Set1') +
-        labs(y = 'log likelihood ratio\n←opposition                                 support→') +
+        scale_color_brewer(palette = 'Set1', 
+                           guide = 'none') +
+        labs(y = 'log likelihood ratio\n←support                                 oppose→') +
         ggtitle(title)
 }
 
 p_df |> 
     map(llr) |> 
     imap(llr_by_noun_plot) |> 
-    wrap_plots(nrow = 1, guides = 'collect')
+    wrap_plots(nrow = 1, guides = 'collect') +
+    plot_layout(axis_titles = 'collect')
+
 ## TODO: feed in the combined plot
 plotly::ggplotly()
 

### 2024-02-24 R/load_coding.R
track stuff that hadn't been tracked

@@ -1,0 +1,43 @@
+## Load the manual codes
+load_manual_coding = function(filter_na = FALSE) {
+    dataf = read_rds(here(data_dir, '07_coding.Rds')) |> 
+        select(-starts_with('support'), 
+               -starts_with('notes')) |> 
+        mutate(support = case_when(
+            mode == 'y' ~ 'support', 
+            mode == 'n' ~ 'oppose', 
+            TRUE ~ NA_character_
+        )) |> 
+        select(-mode)
+    if (filter_na) {
+        return(filter(dataf, !is.na(support)))
+    } else {
+        return(dataf)
+    }
+}
+
+## Load the imputed codes for the full corpus
+load_imputed_full = function() {
+    here(data_dir, '08_imputed.parquet') |> 
+        open_dataset() |> 
+        rename(support = .pred_class) |> 
+        select(-text)
+}
+
+## Load the imputed codes, filtering by a confidence threshold
+load_imputed_fltd = function(threshold = .8) {
+    load_imputed_full() |> 
+        filter(.pred_oppose < 1 - threshold | .pred_oppose > threshold)
+}
+
+## Helper to map a function over the three datasets and return a combined dataframe
+mapped = function(input, f) {
+    input |> 
+        map(collect) |> 
+        map(f) |> 
+        bind_rows(.id = 'coding')
+}
+mapped2 = function(input1, input2, f) {
+    map2(input1, input2, f) |> 
+        bind_rows(.id = 'coding')
+}

### 2024-02-24 reg.gov.api/R/extract_text.R

@@ -1,0 +1,95 @@
+#' Extract text from a PDF or DOCX to a TXT file
+#'
+#' @param source Source file
+#' @param target Target TXT file; if NULL does not write output
+#' @param force  Extract text even if target file already exists
+#' @param ... Other arguments, passed down to `extract_text_pdf()` or `extract_text_docx()`
+#' @return Extracted text (invisible)
+#' @export
+extract_text = function(source,
+                        target = NULL,
+                        force = FALSE,
+                        verbose = FALSE,
+                        ...) {
+    if (verbose) {
+        message(source)
+    }
+    type = tools::file_ext(source)
+    valid_types = c('pdf', 'docx')
+    assertthat::assert_that(type %in% valid_types,
+                            msg = 'Source document must be PDF or DOCX')
+
+    if (is.null(target) || !file.exists(target) || force) {
+        if (type == 'pdf') {
+            # message('pdf branch')
+            text = extract_text_pdf(source, ...)
+        } else if (type == 'docx') {
+            # message('docx branch')
+            text = extract_text_docx(source, ...)
+        } else {
+            stop('Unhandled document type in extract_text()')
+        }
+
+        if (!is.null(target)) {
+            readr::write_lines(text, target)
+        }
+    } else {
+        if (verbose) {
+            message(target, ' exists')
+        }
+        text = readr::read_lines(target)
+    }
+    return(invisible(text))
+}
+
+#' Extract text from a PDF file
+#'
+#' @param source Source file to extract test from
+#' @param ocr_threshold If text length in characters is less than this, run it through OCR
+#' @return Character vector containing the extracted text
+extract_text_pdf = function(source,
+                            ocr_threshold = 50) {
+    text = pdftools::pdf_text(source) %>%
+        stringr::str_flatten(collapse = '\n')
+
+    check_text = text %>%
+        stringr::str_flatten() %>%
+        stringr::str_squish()
+
+    ## If the text is too short, run it through OCR
+    if (stringr::str_length(check_text) < ocr_threshold) {
+        image_files = source %>%
+            stringr::str_remove('.pdf') %>%
+            stringr::str_c('-', 1:pdftools::pdf_info(source)$pages) %>%
+            stringr::str_c('.png')
+        pdftools::pdf_convert(source,
+                              format = 'png',
+                              dpi = 150,
+                              filenames = image_files)
[…]

### 2024-02-24 scripts/10_keyword_occurrence.R

@@ -1,0 +1,48 @@
+library(tidyverse)
+theme_set(theme_minimal())
+library(ggforce)
+library(patchwork)
+library(here)
+library(arrow)
+library(tictoc)
+
+extract_noun = function(df, in_col = bigram, out_col = noun) {
+    mutate(df, {{ out_col }} := str_split_i({{ in_col }}, '_', -1))
+}
+
+## Load data ----
+data_dir = here('data')
+source(here('R', 'load_coding.R'))
+
+coded = list(manual = load_manual_coding(filter_na = TRUE), 
+             filtered = load_imputed_fltd(), 
+             imputed = load_imputed_full())
+
+## Analyze keyword occurrence ----
+## TODO: use lookaheads/lookbacks to avoid catching "Strengthening Transparency in Regulatory Science" and "Environmental Protection Agency"
+kword_occurence = function(coded_df) {
+    open_dataset(here('data', '03_text.parquet')) |>
+        inner_join(coded_df, by = 'comment_id') |> 
+        ## <https://github.com/apache/arrow/issues/40220>
+        collect() |>
+        mutate(science =     str_detect(text, regex('scien', ignore_case = TRUE)), 
+               health =      str_detect(text, regex('health|medic', ignore_case = TRUE)), 
+               environment = str_detect(text, regex('(?!(?-i)Environmental Protection Agency)(enviro|ecolog)\\w+', ignore_case = TRUE)), 
+               economy =     str_detect(text, regex('business|econom', ignore_case = TRUE)), 
+               regulation =  str_detect(text, regex('(?!(?-i)Regulatory Science)(regulat\\w*)', ignore_case = TRUE))) |> 
+        group_by(support, comment_id) |> 
+        summarize(across(c(science, health, 
+                           environment, economy, regulation), 
+                         any)) |> 
+        ungroup() |> 
+        group_by(support) |> 
+        summarize(n = n(),
+                  across(c(science, health, 
+                           environment, economy, regulation), 
+                         mean)) |> 
+        arrange(support)
+}
+map(coded, kword_occurence) |> 
+    map(collect) |> 
+    bind_rows(.id = 'coding')
+

### 2024-02-24 scripts/10_synonyms.R
word2vec variant to check keywords

@@ -1,0 +1,130 @@
+## <https://juliasilge.com/blog/tidy-word-vectors/>
+library(tidyverse)
+library(tidytext)
+library(widyr)
+library(irlba)
+
+library(arrow)
+library(here)
+
+word_vec_file = here('data', '10_word_vec.Rds')
+
+## SVD step is rather slow, so we'll only rebuild the model if necessary
+if (!file.exists(word_vec_file)) {
+    ## Load data ----
+    ## <https://universaldependencies.org/u/pos/>
+    ann = open_dataset(here('data', '03_annotations')) |> 
+        filter(pos %in% c('ADJ', 'NOUN', 'VERB', 'ADV')) |> 
+        mutate(token = str_c(lemma, '_', pos)) |> 
+        select(comment_id, doc_id, sentence_id, token)
+    
+    ## Unigram probabilities ----
+    unigram_p = ann |> 
+        count(token) |> 
+        collect() |> 
+        mutate(p = n / sum(n))
+    
+    ## Co-occurrence probabilities and PMI ---
+    coocc_p = ann |> 
+        ## Construct unique sentence IDs
+        mutate(context_id = str_c(doc_id, '-', sentence_id)) |> 
+        select(context_id, token) |> 
+        collect() |> 
+        pairwise_count(token, context_id, 
+                       diag = TRUE, sort = TRUE) |> 
+        filter(n > 20) |> 
+        mutate(p = n / sum(n)) |> 
+        as_arrow_table()
+    
+    pmi = coocc_p |> 
+        left_join(unigram_p, 
+                  by = join_by(item1 == token), 
+                  suffix = c('', '1')) |> 
+        left_join(unigram_p, 
+                  by = join_by(item2 == token), 
+                  suffix = c('', '2')) |> 
+        mutate(pmi = log10(p / p1 / p2)) |> 
+        select(item1, item2, p, p1, p2, pmi)
+    
+    # pmi |> 
+    #     filter(item1 == 'science_NOUN', 
+    #            str_detect(item2, '_NOUN')) |> 
+    #     collect()
+    # 
+    # pmi |> 
+    #     filter(item1 == 'environment_NOUN') |> 
+    #     collect()
+    
+    ## SVD ----
+    pmi_sparse = pmi |> 
+        collect() |> 
+        cast_sparse(item1, item2, pmi)
+    
[…]

### 2024-03-27 paper/methods.qmd
table/plot output

@@
 After scraping, our corpus included the text of 22,390 comments and attachments for 20,787 distinct comments.
-*[try reruning scraper sometime]*  By number of attachments, the median document had no attachments and the maximum number of attachments was 17.
+*[try reruning scraper sometime]*  We refer to this collection of comments and attachments as the *ST corpus*.
+By number of attachments, the median document had no attachments and the maximum number of attachments was 17.
 
-## Sampling
+## Manual coding
 
@@
 
-Further exploration — and a fortuitous coding error — identified a solution.
+### Sampling
+
+Further exploration — and a serendipitous coding error — identified a solution to the oversampling challenge.
 Using the `spaCy` natural language processing (NLP) suite and its R API `spacyR` *[versions]*, we analyzed the text of each document into individual words ("tokens") and classified each term's part of speech (noun, adjective, etc.).
@@
 
-## Manual coding
+### Manual coding
 
 At this point, [A1] brought the other collaborators into the project.
-Without informing the other authors of the research question, all authors independently coded each comment in the sample as supporting or opposing ST, or too ambiguous to classify.
-Krippendorff's alpha was used to assess inter-rater reliability *[https://link.springer.com/article/10.1007/s11135-004-8107-7]*; this statistic ranges from -1 to 1, with 1 indicating perfect agreement across all raters and 0 indicating "no agreement beyond chance" *[https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4974794/]*.
+Without informing the other authors of the research question, all authors independently coded each comment in the sample as supporting or opposing ST, or too ambiguous to classify.^[Among qualitative social scientists, this is referred to as "deductive coding." "Inductive" or "open" coding, by contrast, does not start with pre-defined codes, and is about developing a novel coding scheme, in a rather Baconian sense of induction.]  Krippendorff's alpha was used to assess inter-rater reliability *[https://link.springer.com/article/10.1007/s11135-004-8107-7]*; this statistic ranges from -1 to 1, with 1 indicating perfect agreement across all raters and 0 indicating "no agreement beyond chance" *[https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4974794/]*.
 After an initial round of independent coding, Krippendorff's alpha was 0.86, with discordance on 74 / 814 comments.
@@
 
-## Imputation
+### Imputation
 
@@
 
+## Keyword synonyms and exclusions
+
+Keyword detection is vulnerable to both false negatives (ignoring synonyms or closely-related concepts that do not strictly match chosen keywords) and false positives (counting unrelated concepts that do match chosen keywords).
+For example, use of the rule's title, "Strengthening Transparency in Regulatory Science," should not count as evoking or referring to science.
+
+To analyze the corpus for these possibilities, we implemented a lightweight variation on word embeddings.
+Word embeddings attempt to represent words (or whatever unit of text) as vectors in an abstract space such that two vectors point in the same direction (more formally: have high cosine similarity) insofar as the corresponding words are semantically similar.
+While word embeddings are usually created using neural network techniques, *[<https://proceedings.neurips.cc/paper/2014/hash/feab05aa91085b7a8012516bc3533958-Abstract.html>]* found that a less computationally-intensive method (partial singular value decomposition [SVD] of a pointwise mutual information matrix) produced similar results.
+We implemented this method in R, following the example of *[Silge blogpost]*.
+
+Specifically, we first filtered the spaCy annotated unigrams (from 20,787 comments) to those that were tagged as adjectives, nouns, verbs, and adverbs, since we judged these to be the most likely to contain keyword matches.
+Tokens were defined as a combination of a word lemma (standardized word form) and part of speech tag, such as `scientifically_ADV`.
+This resulted in 78,744 tokens.
+Marginal occurrence probabilities (across the entire corpus annotated with spaCy) were calculated for each token.
+Two tokens were determined to co-occur if they occurred in the same sentence (as identified by spaCy) in the same comment/attachment.
+Co-occurrences were filtered at 20; that is, if a pair of tokens had 20 or fewer co-occurrences, this was treated as 0 co-occurrences.
+After filtering, 8,364 tokens had at least one non-zero occurrence.
+Co-occurrence / joint probabilities were calculated, and these were combined with marginal probabilities to calculate pointwise mutual information (pmi):
+
+\[ pmi(x, y) = \log \frac{p(x, y)}{p(x) p(y) }\]
+
+We then calculated the partial SVD of this pmi matrix using the `irlba` package *[version]*, selecting 256 dimensions.
+This resulted in a 8,364 $\times$ 256 embedding matrix, with rows representing tokens and columns representing dimensions of the embedding space.
+
+Given a keyword search string, such as `scien`, we first searched the embedding vocabulary (8,364 tokens) for matches.
+For each match, identifying similar tokens was simply a matter of matrix multiplication and sorting.
+The sorted results were filtered to have the same part of speech as the match, and then the top 5 results were displayed for each match.
+We perused the resulting list to identify potential synonyms and matches that needed to be excluded.
+Some additional, multiword false positive matches were identified based on our familiarity with the corpus, such as "Environmental Protection Agency" or the title of ST.
+
 ## Text analysis
@@
 Text analysis was conducted as reported in the main text, with code written in R *[version]* and making extensive use of the `tidyverse` suite *[version]*.
+
+## Inferential statistics and inductive scope
+
+While the ST corpus can be considered a sample of US public opinion on ST, almost all comments come from a relatively highly engaged and haphazard subset:  people on email lists for environmental advocacy organizations, regulated industries, state regulators, readers of anti-environmentalist blogs.
+It therefore cannot be considered a representative sample of anyone except "the kinds of people who were likely to have submitted comments on ST."  And for this population the ST corpus is, as it were, a nearly complete sample:  insofar as someone was likely to have submitted comments on ST, we probably have a comment from them in the ST corpus.
+
+Thus, we believe that so-called inferential statistics — things like p-values from statistical hypothesis tests and confidence intervals — would be misleading, as they would incorrectly indicate that we can formally generalize from the ST corpus to some more-or-less well-defined larger population.
+We therefore did not calculate, and do not report here, any such statistics.
 

### 2024-03-27 R/extract_noun.R

@@ -1,0 +1,3 @@
+extract_noun = function(df, in_col = bigram, out_col = noun) {
+    mutate(df, {{ out_col }} := str_split_i({{ in_col }}, '_', -1))
+}

### 2024-03-27 R/load_coding.R

@@ -39,5 +39,6 @@ mapped = function(input, f) {
 }
 mapped2 = function(input1, input2, f) {
     map2(input1, input2, f) |> 
-        bind_rows(.id = 'coding')
+        bind_rows(.id = 'coding') |> 
+        mutate(coding = fct_inorder(coding))
 }

### 2024-03-27 R/tab_out.R

@@ -1,0 +1,6 @@
+tab_out = function(gt_tab, filename) {
+    gtsave(gt_tab, here(out_dir, glue('{filename}.png')))
+    write_rds(gt_tab, here(out_dir, glue('{filename}.Rds')))
+}
+## tab_out(n_docs_tab, '09_n_docs')
+

### 2024-03-27 scripts/09_bigram_analysis.R

@@ -1,18 +1,22 @@
 library(tidyverse)
-theme_set(theme_minimal())
+
+theme_set(theme_bw())
 library(ggforce)
 library(patchwork)
+library(gt)
+
 library(here)
 library(arrow)
+library(glue)
 
-extract_noun = function(df, in_col = bigram, out_col = noun) {
-    mutate(df, {{ out_col }} := str_split_i({{ in_col }}, '_', -1))
-}
+source(here('R', 'tab_out.R'))
+source(here('R', 'extract_noun.R'))
 
 ## TODO: write out plots throughout
 
 ## Load data ----
 data_dir = here('data')
+out_dir = here('out')
 source(here('R', 'load_coding.R'))
 
 coded = list(manual = load_manual_coding(), 
@@ -20,16 +24,6 @@ coded = list(manual = load_manual_coding(),
              imputed = load_imputed_full())
 
 map(coded, nrow)
-
-## Number of docs by support
-## Here we only want number of docs *w/ bigrams*
-# count_docs = function(coded_df) {
-#     count(coded_df, support, name = 'n_docs')
-# }
-# # n_docs = count_docs(coding)
-# n_docs = map(coded, count_docs)
-# 
-# mapped(n_docs, ~ mutate(.x, share = n_docs / sum(n_docs)))
 
 
 ## Load bigrams ----
@@ -53,30 +47,96 @@ n_docs = map(bigrams, ~ {.x |>
         distinct(comment_id, support) |> 
         count(support, name = 'n_docs')})
 
+## table: number of docs by support
+n_docs_tab = bind_rows(n_docs, .id = 'dataset') |> 
+    group_by(dataset) |> 
+    mutate(perc = n_docs / sum(n_docs)) |> 
+    ungroup() |> 
+    pivot_wider(names_from = support, values_from = c(n_docs, perc)) |> 
+    rowwise() |> 
+    mutate(total = sum(c_across(starts_with('n_docs')))) |> 
+    ungroup() |> 
+    gt(rowname_col = 'dataset') |> 
+    fmt_percent(starts_with('perc'), decimals = 0) |> 
+    fmt_integer(c(starts_with('n_doc'), total)) |> 
+    cols_merge(columns = ends_with('oppose'), 
+               pattern = '{1} ({2})') |> 
+    cols_merge(columns = ends_with('support'), 
+               pattern = '{1} ({2})') |> 
+    cols_label(n_docs_oppose ~ 'oppose', 
+               n_docs_support ~ 'support') |> 
[…]

### 2024-03-27 scripts/10_synonyms.R

@@ -18,11 +18,19 @@ if (!file.exists(word_vec_file)) {
         mutate(token = str_c(lemma, '_', pos)) |> 
         select(comment_id, doc_id, sentence_id, token)
     
+    ## 20,787 comments
+    ann |> 
+        pull(comment_id) |> 
+        n_distinct()
+    
     ## Unigram probabilities ----
     unigram_p = ann |> 
         count(token) |> 
         collect() |> 
         mutate(p = n / sum(n))
+    
+    ## 78,744
+    nrow(unigram_p)
     
     ## Co-occurrence probabilities and PMI ---
     coocc_p = ann |> 
@@ -35,6 +43,11 @@ if (!file.exists(word_vec_file)) {
         filter(n > 20) |> 
         mutate(p = n / sum(n)) |> 
         as_arrow_table()
+    
+    ## 8,364
+    c(pull(coocc_p, item1), 
+      pull(coocc_p, item2)) |> 
+        n_distinct()
     
     pmi = coocc_p |> 
         left_join(unigram_p, 
@@ -104,7 +117,7 @@ get_variants = function(pattern, n = 10, filter_pos = TRUE) {
 ## Actually do the searching ----
 ## Science: 'scien'
 ## no synonyms that need to be included
-## exclude: conscience, epascientificresearchtransperancyplan.pdf
+## exclude: conscience, epascientificresearchtransperancyplan.pdf, Strengthening Transparency in Regulatory Science
 ## scientifically_ADV is most similar to right_ADV, correctly_ADV, importantly_ADV
 get_variants('scien', n = 5) |> view(title = 'scien')
 

### 2024-03-27 scripts/11_keyword_occurrence.R

@@ -2,13 +2,16 @@ library(tidyverse)
 theme_set(theme_minimal())
 library(ggforce)
 library(patchwork)
+library(gt)
+
 library(here)
 library(arrow)
+library(glue)
 library(tictoc)
 
-extract_noun = function(df, in_col = bigram, out_col = noun) {
-    mutate(df, {{ out_col }} := str_split_i({{ in_col }}, '_', -1))
-}
+source(here('R', 'extract_noun.R'))
+source(here('R', 'tab_out.R'))
+out_dir = here('out')
 
 ## Load data ----
 data_dir = here('data')
@@ -19,17 +22,26 @@ coded = list(manual = load_manual_coding(filter_na = TRUE),
              imputed = load_imputed_full())
 
 ## Analyze keyword occurrence ----
-## TODO: use lookaheads/lookbacks to avoid catching "Strengthening Transparency in Regulatory Science" and "Environmental Protection Agency"
+## TODO: use lookaheads/lookbacks to avoid catching "Strengthening Transparency in Regulatory Science"
 kword_occurence = function(coded_df) {
     open_dataset(here('data', '03_text.parquet')) |>
         inner_join(coded_df, by = 'comment_id') |> 
         ## <https://github.com/apache/arrow/issues/40220>
         collect() |>
-        mutate(science =     str_detect(text, regex('scien', ignore_case = TRUE)), 
-               health =      str_detect(text, regex('health|medic', ignore_case = TRUE)), 
-               environment = str_detect(text, regex('(?!(?-i)Environmental Protection Agency)(enviro|ecolog)\\w+', ignore_case = TRUE)), 
-               economy =     str_detect(text, regex('business|econom', ignore_case = TRUE)), 
-               regulation =  str_detect(text, regex('(?!(?-i)Regulatory Science)(regulat\\w*)', ignore_case = TRUE))) |> 
+        mutate(science = str_detect(text, 
+                                    regex('(?!(?-i)Regulatory Science)(science)', 
+                                          ignore_case = TRUE)), 
+               health = str_detect(text, regex('health|medic', 
+                                               ignore_case = TRUE)), 
+               environment = str_detect(text, 
+                                        regex('(?!(?-i)Environmental Protection Agency)(enviro|ecolog)', 
+                                              ignore_case = TRUE)), 
+               economy = str_detect(text, 
+                                    regex('business|econom', 
[…]

### 2024-05-15 paper/paper.qmd
initial work on paper

---
title: "The aims of science and an open science controversy at USEPA"
authors:
  - name: "Daniel J. Hicks"
    email: "dhicks4@ucmerced.edu"
    affiliations: 
      - id: ucm
        name: "University of California, Merced"
        city: "Merced"
        country: "CA, USA"
    orcid: "0000-0001-7945-4416"
    corresponding: true
  - name: "Emilio J. C. Lobato"
    affiliations:
      - ref: ucm
    orcid: "*[todo]*"
  - name: "Cosmo Campbell"
    affiliations:
      - ref: ucm
  - name: "Joseph Dad"
    affiliations:
      - ref: ucm

abstract: "abstract"

acknowledgments: "acknowledgments"

format:
  html:
    toc: true
    number-sections: true
  pdf:
    toc: true
    number-sections: true
    link-citations: true
    pdf-engine: "lualatex"
    mainfont: "Times New Roman"
    fontsize: "12pt"
    geometry: 
      - "margin = 1in"
    linestretch: 1.5

crossref:
  fig-prefix: ""
  tbl-prefix: ""

execute:
  eval: true
  echo: false
  warning: true
  message: false
---

<!-- 
BJPS submission guidelines: 
<https://www.journals.uchicago.edu/journals/bjps/instruct>

"strict 24 page limit," though not including references
-->

```{r}
library(readr)
library(gt)
library(here)

out_dir = here('out')
```

# Introduction

*[todo]*

# Methods overview

*[todo]*
Details of methods are discussed in @sec-methods.

# Results

Table @tbl-n-docs shows the number of comments in the corpus, across the three coding methods and by support or opposition to ST; table @tbl-n-docs-1 shows counts for the whole corpus, while table @tbl-n-docs-2 shows counts for comments containing science or health bigrams.  In the whole corpus, opposing comments substantially outnumbered supporting comments, by a ratio of 3- or 4-to-1.  Also in the whole corpus, the ML classifier had a somewhat higher rate of supporting comments (~27%) than we identified with manual coding (19%).  In the filtered and imputed coding, most supporting comments did not use science or health bigrams.

```{r}
#| eval: true
#| label: "tbl-n-docs"
#| tbl-cap: "Document counts, by coding method and opposition/support"
#| tbl-subcap:
#|   - "Whole corpus"
#|   - "Comments with science/health bigrams"
#| layout-cols: 2
read_rds(here(out_dir, '09_n_docs.Rds'))
read_rds(here(out_dir, '09_n_docs_bigrams.Rds'))
```

In the manual coding, the 33 "neither" comments were almost all procedural requests to extend the public comment period, with a few comments discussing an issue that was brought up as an example in the first NPR (the linear no-threshold model for radiation hazard) that we did not consider relevant to our analysis.  Because the ML classifier was not trained on these 33 "neither" comments, it forced all such comments into either the "support" or the "oppose" category.  Notably, the comments that were dropped by the "low confidence" filter (the "filtered" coding) are generally procedural requests to extend the public comment period.

## Bigram analysis

Figure @fig-bigram-occurrence shows the 15 most common bigrams (both science and health) by ST support (rows of panels) across the three codings (columns of panels). Three bigrams occur in the majority of opposing comments, across all codings: "public health," "available science" (short for "best available science"), and "good science." Other common bigrams used by opponents include "human health," "environmental health," "reliable science," and "medical science."

![15 most common adjective-noun bigrams, by support for ST (rows of panels) and codings (columns of panels). Science bigrams in blue, health bigrams in red.  Relative document frequency is based only on comments containing science/health bigrams.](../out/09_occurrence.png){#fig-bigram-occurrence}

Among supporting comments that use bigrams, they are less consistent in which bigrams they use:  no bigrams appear in the majority of comments.  The most common bigram is "secret science," followed by "regulatory science," "good science," and "public health"; though these three occur in fewer than 25% of comments.

There are also notable differences in the overall tone of the adjectives in supporters' and opponents' top bigrams.  Supporters seem to express skepticism or cynicism about the quality of scientific research: science is described as secret, secretive, phony, false, fake, shady, or deceptive.  In the context of these skeptical assessments, seemingly-positive valence bigrams — such as real, true, verifiable, and sound science — might be used as privatives, that is, expressing the idea that actually-existing science is *not* real, true, verifiable, or sound.  A skeptical or cynical assessment of science would provide support for ST insofar as open data would make science less secretive and phony and more verifiable and true, etc.  @HicksOpenScienceReplication2023 argued that there is no empirical evidence of a replication crisis in environmental epidemiology [though see @BagiletEstimatingSmallEffects], and that ST's open data requirements would not mitigate widespread irreplicability even if it were a problem in that field.

Table @tbl-noun-occ and figure @fig-noun-occ examine the occurrence of bigrams, aggregated by noun, for supporters and opponents.  For both supporters and opponents, almost all comments (~90% or more) that use bigrams use at least one science bigram.  For health bigrams and opponents, this is similar: almost all opponents use at least one health bigram.  But health bigrams are rare among supporters, with at most 26% of supportive comments (that use bigrams) using at least one health bigram.

```{r}
#| eval: true
#| label: "tbl-noun-occ"
#| tbl-cap: "Bigram noun occurrence, by coding method and noun"
read_rds(here(out_dir, '09_noun_occ_tab.Rds'))
```

![Bigram noun occurrence, by coding method and noun, only comments containing science/health bigrams. Vertical lines connect the two nouns for a given coding and support.](../out/09_noun_occurrence.png){#fig-noun-occ}

This analysis indicates that supporters put much more emphasis on science than health in their comments on ST; while opponents emphasize science and health roughly equally.

## Keyword analysis

The bigram analysis is limited to documents that contain adjective-noun bigrams, as identified by the NLP parser.  To expand this analysis, we switch to a keyword search of the entire corpus.  Table @tbl-keyword shows the occurrence of hits for five keyword searches, corresponding to science, health/medicine, the environment and ecology, business and the economy, and regulation.

```{r}
#| eval: true
#| label: "tbl-keyword"
#| tbl-cap: "Keyword occurrence, by coding method and support. Footnotes indicate regular expressions used for search."
read_rds(here(out_dir, '11_keyword_tab.Rds'))
```

In the bigram analysis, both supporters and opponents almost always used at least one science bigram.  This is still the case in the manually coding; but the rate is somewhat lower for supporters in the two imputed codings, ~65%.  Opponents are also likely to use health (~95%), environment (77% or more), and regulation (~70%) keywords.  By contrast, supporters are unlikely to use any keywords except science (and a slight majority for regulation keywords in the manual coding).

We were surprised that supporters were unlikely to use business or regulation keywords, since we expected that they might support ST on the grounds that it would reduce regulation and thereby promote economic growth or business development.  To get a better sense of how supporters of ST were commenting, we calculated the log likelihood ratio for supporters and opponents for all (cleaned and lemmatized) unigram tokens in the corpus; see figure @fig-token-llr.  This use of log likelihood ratio identifies distinctive or characteristic terms, which have a relatively high likelihood of appearing on one side and a relatively low likelihood on the other.

![15 most distinctive terms (unigrams) for supporters and opponents, by coding. Terms in the top half are distinctive to opponents of ST; terms in the bottom half are distinctive to supporters.](../out/11_llr.png){#fig-token-llr}

The magnitude of log likelihood ratios here reinforces the finding, from the bigram analysis, that supporters are less consistent in their language than opponents.  The most distinctive terms among supporters appear to correspond to the phrase "please stop/end/show your secret science," variations of which were common in the manual coding.  The tokens "regulation" (manual), "burdensome" (filtered, imputed), "regulate" (imputed), and "small" (all three) might correspond to concerns that burdensome regulations harm small businesses.  For opponents, the most distinctive term in all three codings is "health," and "protect" is second or third.

These distinctive terms indicate not just what each side is saying about ST, but also what the other side is *not* saying.  Opponents do not deploy the "secret science" framing used by supporters, and supporters do not talk about USEPA's mission to protect human health and the environment.

# Discussion

## Two views of the aims of science

We argue that the results above can be fruitfully understood in terms of *[Hicks']* account of two views of the aims of science.  Specifically, we claim that supporters deployed the narrow view, while opponents deployed the broad view.  On the narrow view of supporters of ST, science has only epistemic aims.  The practical applications of science are at best irrelevant to the standards for good scientific research, and allowing such applications to shape things like data collection and analysis is comparable to — indeed, *is* — research misconduct.  The narrow view's exclusive focus on epistemic aims provides argumentative support for truth-promoting practices, such as those promoted by the open science movement, wherever it is feasible to do so.  Deployment of the narrow view explains why supporters said a great deal about science, but almost nothing about public health and the environment, and even very little about business and the economy.  All of these aspects of environmental policy are irrelevant to the standards for good scientific research.

By contrast, opponents deployed the broad view, on which science has both epistemic and practical aims.  Specifically, they appealed to the practical aims of protecting human health and the environment, raising concerns that ST's open data requirements would frustrate the pursuit of these aims, and was therefore "anti-science" *[Hicks]*.  In this way, the broad view provides an argumentative challenge to ST.  Deployment of the broad view explains why opponents talked about public health and the environment almost as frequently as science itself.

Deployment of the broad view might also explain why opponents said relatively little about business and the economy.  On Hicks' theory of value, protecting human health and the environment are aims of environmental public health, but the impacts of regulation on the economy are still accidents, and thus irrelevant to the standards of good scientific research.  @HicksValuesDisclosuresTrust2022 provide survey evidence that, in the context of chemical regulation, members of the public put more trust in a scientist who discloses public health values than one who does not disclose values or discloses economic values.

As an alternative to the aims approach, our findings might be interpreted in terms of inductive risk, and specifically differences between supporters and opponents in the relative importance of over- vs. under-regulation *[cites]*.  That is, supporters of ST might put more weight on over-regulation and preventable economic harms; while opponents of ST might put more weight on under-regulation and preventable disease and loss of life due to pollution.

An inductive risk framing might work well for opponents, who are very likely to talk about public health, the environment, and ecology.  But it does not seem to fit supporters, who talk about the economy about as often as they talk about the environment, and are less likely than not to talk about regulation.  We therefore take the aims approach to be a better fit for analyzing this particular controversy.

## The views as interpretive frames, not beliefs

## The role of advocacy organizations

*[Starbird: participatory disinformation]*

\appendix

# Methods {#sec-methods}

{{< include methods.qmd >}}

### 2024-05-15 scripts/09_bigram_analysis.R

@@ -12,7 +12,38 @@ library(glue)
 source(here('R', 'tab_out.R'))
 source(here('R', 'extract_noun.R'))
 
-## TODO: write out plots throughout
+build_n_docs_tab = function(dl, has_missing = FALSE) {
+    dl |> 
+        map(~ {.x |> 
+                distinct(comment_id, support) |> 
+                count(support, name = 'n_docs')}) |> 
+        map(collect) |> 
+        bind_rows(.id = 'coding') |> 
+        group_by(coding) |> 
+        mutate(perc = n_docs / sum(n_docs)) |> 
+        ungroup() |> 
+        pivot_wider(names_from = support, 
+                    values_from = c(n_docs, perc), 
+                    values_fill = 0L) |> 
+        rowwise() |> 
+        mutate(total = sum(c_across(starts_with('n_docs')))) |> 
+        ungroup() |> 
+        gt(rowname_col = 'coding') |> 
+        fmt_percent(starts_with('perc'), decimals = 0) |> 
+        fmt_integer(c(starts_with('n_doc'), total)) |> 
+        cols_merge(columns = ends_with('oppose'), 
+                   pattern = '{1} ({2})') |> 
+        cols_merge(columns = ends_with('support'), 
+                   pattern = '{1} ({2})') %>% 
+        {if (has_missing) {
+            cols_merge(., 
+                       columns = any_of(ends_with('NA')),
+                       pattern = '{1} ({2})')}
+            else {identity(.)}} |>
+        cols_label(n_docs_oppose ~ 'oppose',
+                   n_docs_support ~ 'support',
+                   any_of('n_docs_NA') ~ 'neither')
+}
 
 ## Load data ----
 data_dir = here('data')
@@ -25,6 +56,11 @@ coded = list(manual = load_manual_coding(),
 
 map(coded, nrow)
 
+
+n_docs_tab = build_n_docs_tab(coded, 
+                              has_missing = TRUE)
+n_docs_tab
+tab_out(n_docs_tab, '09_n_docs')
 
 ## Load bigrams ----
 load_bigrams = function(coded_df) {
@@ -43,31 +79,9 @@ bigrams |>
             pull(comment_id) |> 
             n_distinct()})
 
-n_docs = map(bigrams, ~ {.x |> 
-        distinct(comment_id, support) |> 
-        count(support, name = 'n_docs')})
-
[…]

### 2024-05-15 scripts/11_keyword_occurrence.R

@@ -3,6 +3,7 @@ theme_set(theme_minimal())
 library(ggforce)
 library(patchwork)
 library(gt)
+library(tidytext)
 
 library(here)
 library(arrow)
@@ -22,14 +23,13 @@ coded = list(manual = load_manual_coding(filter_na = TRUE),
              imputed = load_imputed_full())
 
 ## Analyze keyword occurrence ----
-## TODO: use lookaheads/lookbacks to avoid catching "Strengthening Transparency in Regulatory Science"
 kword_occurence = function(coded_df) {
     open_dataset(here('data', '03_text.parquet')) |>
         inner_join(coded_df, by = 'comment_id') |> 
         ## <https://github.com/apache/arrow/issues/40220>
         collect() |>
         mutate(science = str_detect(text, 
-                                    regex('(?!(?-i)Regulatory Science)(science)', 
+                                    regex('(?!(?-i)Regulatory Science)(scien)', 
                                           ignore_case = TRUE)), 
                health = str_detect(text, regex('health|medic', 
                                                ignore_case = TRUE)), 
@@ -37,7 +37,7 @@ kword_occurence = function(coded_df) {
                                         regex('(?!(?-i)Environmental Protection Agency)(enviro|ecolog)', 
                                               ignore_case = TRUE)), 
                economy = str_detect(text, 
-                                    regex('business|econom', 
+                                    regex('business|econom|industr', 
                                           ignore_case = TRUE)), 
                regulation = str_detect(text, 
                                        regex('(?!(?-i)Regulatory Science)(regulat)', 
@@ -74,13 +74,13 @@ kword_tab = kword_df |>
     cols_merge_n_pct(economy_sum, economy_mean) |> 
     cols_merge_n_pct(regulation_sum, regulation_mean) |> 
     cols_label_with(fn = ~ str_remove(.x, '_sum')) |> 
-    tab_footnote('(?!(?-i)Regulatory Science)(science)', 
+    tab_footnote('(?!(?-i)Regulatory Science)(scien)', 
                  locations = cells_column_labels(c(science_sum))) |> 
     tab_footnote('health|medic', 
                  locations = cells_column_labels(c(health_sum))) |> 
     tab_footnote('(?!(?-i)Environmental Protection Agency)(enviro|ecolog)', 
[…]

### 2024-05-16 paper/methods.qmd
initial work on paper
date and length plots

@@
 After scraping, our corpus included the text of 22,390 comments and attachments for 20,787 distinct comments.
-*[try reruning scraper sometime]*  We refer to this collection of comments and attachments as the *ST corpus*.
+We refer to this collection of comments and attachments as the *ST corpus*.
 By number of attachments, the median document had no attachments and the maximum number of attachments was 17.
+
+Comments were primarily submitted during two public comment periods, one following the publication of the Notice of Proposed Rulemaking (NPR) on *[date]* and the second in response to the Revised Notice of Proposed Rulemaking (RNPR) on *[date]*.
+Comments varied in length dramatically: a few handwritten postcards could not be parsed by OCR and were counted as being 0 words in length, and the longest comment (including the comment text and all attachment text) was 467,917 words long.
+The median comment was 236 words long, and 95% of comments were 495 words long or less.
+See figure @fig-date-ecdf.
+
+![(a) Count of submissions, aggregated by week; note the y-axis uses a square root scale. (b) Cumulative frequency of comment lengths, by number of words.](../out/04_date_ecdf.png){#fig-date-ecdf}
 
@@
 Drawing the most extreme documents from either end resulted in a large number of nearly identical "form letter" comments, most likely collected by advocacy organizations using online forms.
-Using phrases that appeared in these comments and simple text matching, [A1] excluded 8,875 from three different forms from the manual coding sample.
+Using phrases that appeared in these comments and simple text matching, [A1] excluded 8,875 comments matching three different forms from the manual coding sample.
 After these exclusions, drawing the most extreme 200 documents (including ties) from each end resulted in *[actual numbers, pre-fixing html issue]* 214 from the "bottom" (apparently supportive) and 201 from the "top" (apparently opposed).
@@
 
-\[ pmi(x, y) = \log \frac{p(x, y)}{p(x) p(y) }\]
+$$ pmi(x, y) = \log \frac{p(x, y)}{p(x) p(y) }$$
 

### 2024-05-16 scripts/04_eda.R

@@ -1,6 +1,7 @@
 ## EDA of parsed comments
 library(tidyverse)
 theme_set(theme_minimal())
+library(patchwork)
 library(here)
 library(arrow)
 
@@ -64,14 +65,23 @@ filter(tokens, pos == 'SYM') |>
 
 ## Comment posting date ----
 ## Two waves: (1) May-August 2018, (2) March-May 2020
-meta |> 
+date_gg = meta |> 
     mutate(month = round_date(date, unit = 'month'), 
            week = round_date(date, unit = 'week')) |> 
     collect() |> 
     ggplot(aes(week)) +
-    geom_bar() +
-    scale_x_date(date_breaks = '4 months') +
-    scale_y_log10()
+    geom_bar(fill = 'blue', color = 'transparent') +
+    geom_vline(xintercept = ym('2018-01',
+                               '2019-01',
+                               '2020-01'), 
+               alpha = .4) +
+    scale_x_date(date_breaks = '4 months', 
+                 date_minor_breaks = '2 months',
+                 date_labels = '%Y-%m') +
+    scale_y_sqrt() +
+    labs(x = 'date', 
+         y = 'submission count\n(weekly total)')
+date_gg
 
 ## first date was 22 May 2017; final date was 8 Dec 2020
 summary(meta$date)
@@ -84,7 +94,7 @@ meta |>
     inner_join(collect(text), by = c('id' = 'comment_id')) |> 
     pull(text) |> 
     str_trunc(500)
-    
+
 
 ## Distribution of comment length ----
 ## In tokens; dropping punctuation, spaces, symbols, unknown
@@ -92,16 +102,27 @@ dtm = tokens |>
     filter(!!!words_filter) |> 
     count(comment_id, lemma)
 
-dtm |> 
+ecdf_gg = dtm |> 
     group_by(comment_id) |> 
     summarize(n = sum(n)) |> 
     collect() |> 
     ggplot(aes(n)) +
     stat_ecdf() +
-    geom_rug(alpha = .01) +
-    scale_x_log10()
+    geom_rug(alpha = 1) +
+    scale_x_log10(labels = scales::label_log()) +
+    scale_y_continuous(labels = scales::label_percent(accuracy = 1)) +
+    labs(x = 'comment length', 
+         y = 'cumulative frequency')
+ecdf_gg
 
 # plotly::ggplotly()
+
+date_gg + ecdf_gg +
+    plot_layout(ncol = 1) +
[…]

### 2025-03-02 paper/methods.qmd
doc length plot

@@
 
-The US General Services Administration receives and publishes public comments on proposed rules using the website Regulations.gov.
-Commenters may submit comments by online form, email, or physical mail, and electronic formats allow commenters to include a number of attachments.
-Attachments are typically detailed comments or to include journal articles, reports, etc., as supporting evidence.
-
-All agency-published documents and public comments associated with a proposed rule are collected together in a docket, e.g., EPA-HQ-OA-2018-0259 for Strengthening Transparency, available at <https://www.regulations.gov/docket/EPA-HQ-OA-2018-0259>.
 Regulations.gov provides an API (application programming interface) via http, meaning a simple script can be used to automatically retrieve ("scrape") the full contents of a given docket.
@@
 
+*[just use black]*
+
 ![(a) Count of submissions, aggregated by week; note the y-axis uses a square root scale. (b) Cumulative frequency of comment lengths, by number of words.](../out/04_date_ecdf.png){#fig-date-ecdf}
@@
 
-Initial, undirected exploration of the corpus by [A1] indicated that (a) there was a qualitatively clear division between supporters and opponents of ST, (b) commenters offered substantive reasons for their position on ST, but (c) opposing comments were much, much more prevalent than supportive comments.
-The first point meant that manual qualitative coding could be used to identify supporting and opposing comments; combined with the second point, this meant that the corpus was suitable for contrasting the arguments given on either side of the ST controversy.
-But the third point meant that a simple random sample would include very few supporters.
+The coding process is summarized in @fig-workflow.
+
+![Visual summary of the coding workflow. Yellow cells indicate the analysis datasets used in the main text.](workflow.jpg){#fig-workflow}
+
+### Sampling {#sec-sampling}
+
+As discussed in @sec-methods-overview, initial exploration of the comments indicated that we could cleanly distinguish supporters from opponents of ST, but also that a simple random sample would be too unbalanced.
 We therefore needed a way to oversample supporters before manual coding, that is, before classifying them as supporters or opponents.
-
-### Sampling
 
@@
 
-To take advantage of the scale of the ST corpus, we used a machine learning model to impute or estimate support or opposition across *[all]* 20,787 comments, using the `tidymodels` framework *[software versions]*.
+To take advantage of the scale of the ST corpus, we used a machine learning model to impute or estimate support or opposition across *[all]* 20,787 comments, using the `tidymodels` framework *[software versions]* [compare this use of supervised machine learning to the approach sketched by @VaesenPotentialSupervisedMachine2022].
 
@@
 Next, the text was tokenized, stopwords removed *[source]*, and tokens lemmatized.
-The resulting token features were hashed to $1024 = 2^10$ features, with signed binary values (0, +1, -1).
+The resulting token features were hashed to $1024 = 2^10^$ features, with signed binary values (0, +1, -1).
 Zero variance features were removed.
@@
 
-## Keyword synonyms and exclusions
+*[The distribution of document lengths, in words (spaCy-annotated tokens), after coding is shown in fig @fig-len. 11_len.png]*
+
+## Keyword synonyms and exclusions {#sec-embeddings}
 
@@
 Specifically, we first filtered the spaCy annotated unigrams (from 20,787 comments) to those that were tagged as adjectives, nouns, verbs, and adverbs, since we judged these to be the most likely to contain keyword matches.
-Tokens were defined as a combination of a word lemma (standardized word form) and part of speech tag, such as `scientifically_ADV`.
-This resulted in 78,744 tokens.
+Tokens were constructed as a combination of a word lemma (standardized word form) and part of speech tag, such as `scientifically_ADV`.
+This resulted in 78,744 distinct tokens.
 Marginal occurrence probabilities (across the entire corpus annotated with spaCy) were calculated for each token.

### 2025-03-02 paper/paper.qmd
doc length plot

@@
 authors:
-  - name: "Daniel J. Hicks"
+  - name: "D. J. Hicks"
     email: "dhicks4@ucmerced.edu"
@@
 
-*[todo]*
+*[
+- general science-and-values
+- aims approach
+- Strengthening Transparency
+]*
 
-# Methods overview
+Text mining is an emerging method for empirical philosophy of science.
+Broadly, text mining aims to identify patterns in the use of terms across a large collection of documents (called a *corpus*).
+While initially developed by computer scientists, text mining methods have been widely adopted by social scientists [@NelsonComputationalGroundedTheory2020; @StoltzMappingTextsComputational2024] and (more contentiously) humanists, under the umbrella of "digital humanities" [@UnderwoodDistantHorizonsDigital2019].
 
-*[todo]*
-Details of methods are discussed in @sec-methods.
+One of the first applications of text mining to philosophy of science was an analysis of Charles Darwin's reading list [@MurdockExplorationExploitationVictorian2017], which used an information-theoretic measure of "surprise" to model Darwin's use of "exploration" and "exploitation" strategies in his reading.
+Malaterre and collaborators have reflexively applied one prominent text mining method, topic modeling, to philosophy of science corpora — especially journal archives — to trace developments in the field over time [@MalaterreRevisitingThreeDecades2019; @MalaterreWhatThisThing2019; @MalaterreEightJournalsEight2020; @MalaterreEarlyDaysContemporary2022].
+Their most recent work combines machine translation with topic modeling, expanding the scope of their analysis to incorporate multiple languages.
+The text mining case studies in @DeBlockDynamicsScienceComputational2022 deploy a variety of methods, including named-entity recognition (NER) and network analysis [@PenceHowNotFight2022] and keyword search to facilitate deep reading [@GinammiBolzanoKantTraditional2022].
+@VaesenPotentialSupervisedMachine2022 argues that supervised machine learning approach can incorporate philosophical expertise in ways that topic modeling cannot; we independently adopted a similar approach in part of the current study.
 
-# Results
+*[methods pieces by Charles?]*
+
+Almost all of these examples of text mining in philosophy of science use, as their corpus, formal publications by scientists (or other scholars, such as philosophers).
+Such corpora are not just apt — we often want to understand how scientists think about, or at least present, their research practices and findings — but also convenient — at least to those of us who have access to a university research library.
+However, for understanding public scientific controversies, we are often more interested in how publics — various interested groups of people — understand science and its relationship to the controversy.
+
+We propose that public comments on high-profile governmental regulatory actions can be apt and convenient corpora for public scientific controversies.
+In the United States, the Administrative Procedure Act requires federal government agencies to follow a "notice and comment" process before adopting new or modified regulations *[cite]*.
+Agencies are required to publish a Notice of Proposed Rulemaking (NPR), including the draft text of the rule and background information to justify it, and then collect and respond to public comments before finalizing the rule.
+Conveniently for our purposes, most comments are made freely and publicly available.
+High-profile proposed rules can collect large numbers of comments, from a variety of different groups and perspectives on the issue at hand in the rule, making it apt for understanding how different groups might, for instance, deploy different views of the aims of science.
+
+ST had an exceptionally high profile, with nearly one million comments.^[In the text mining results below, the corpus contains "only" about 22,000 comments.
+Regulations.gov distinguishes between "comments posted" and "comments received"; a single comment for the "posted" count can include multiple comments for the "received" count.
+For example, the longest comments in the ST corpus comprise tens of thousands of signed electronic form letters.
+This aggregation would count as one "posted" comment but tens of thousands of "received" comments.]  We use this substantial corpus to examine how supporters and opponents talk about science and health, finding that opponents talk about health and the environment about as often as they talk about science; while supporters talk exclusively about science, with a distinctive emphasis on a cynical or skeptical framing of the epistemic reliability of the science utilized by USEPA.
+We argue that these differences correspond to Hicks' [-@HicksWhenVirtuesAre2022] distinction between broad and narrow views of science.
+
+The bulk of this paper gives a non-technical overview of our text mining methods (@sec-methods-overview) and results (@sec-results).
+@sec-methods provides a more technical discussion of our methods, for readers who are familiar with text mining.
+In @sec-discussion, we first argue that our empirical results can be fruitfully understood in terms of the two views of the aims of science.
+We go on to suggest that, in the context of public discourse, deployment of philosophical ideas should be understood as somewhat opportunistic interpretive frames, rather than authentic expressions of beliefs or deep philosophical commitments.
+Finally, we consider the role of advocacy organizations in shaping the content of public scientific controversies, and consider the applicability of a "participatory" model developed in misinformation studies to the case of ST.
+
+# Methods overview {#sec-methods-overview}
+
+We provide technical details on our methods in @sec-methods.  This section provides a non-technical overview.
+
+The US General Services Administration receives and publishes public comments on proposed rules using the website Regulations.gov (<https://www.regulations.gov>).
+Commenters may submit comments by online form, email, or physical mail, and electronic formats allow commenters to include a number of attachments.
+Attachments are typically detailed comments or to include journal articles, reports, etc., as supporting evidence.
+
+All agency-published documents and public comments associated with a proposed rule are collected together in a docket, e.g., EPA-HQ-OA-2018-0259 for Strengthening Transparency, available at <https://www.regulations.gov/docket/EPA-HQ-OA-2018-0259>.
+We collected all public comments from this docket in 2020 and 2021, with a final check several months after ST was finalized (in the last few weeks of the Trump administration) and vacated (a few weeks into the Biden administration).
+
+Initial, undirected exploration of the corpus in 2020-2021 by [A1] indicated that (a) there was a qualitatively clear division between supporters and opponents of ST, (b) commenters offered substantive reasons for their position on ST, but (c) opposing comments were much, much more prevalent than supportive comments.
+The first point meant that manual qualitative coding could be used to identify supporting and opposing comments; combined with the second point, this meant that the corpus was suitable for contrasting the arguments given on either side of the ST controversy.
+But the third point meant that a simple random sample would include very few supporters.
+@sec-sampling details the sampling procedure that we used for the manual coding step.
+
+We used a natural language processing (NLP) package for part-of-speech tagging, specifically identifying nouns and adjectives in the comment and attachment text.
+(From this point forward, attachments were treated as continuations of comment text, and in what follows references to "comments" include the attachments.)  This let us construct adjective-noun bigrams, such as "public health" or "regulatory science."  Focusing on bigrams based on the nouns "health" and "science" let us greatly reduce the size of the vocabulary — the distinct terms used in a text mining analysis — which in turn greatly reduced the computation time necessary for any individual step.
+[A1] had also identified the relationship between science and health, and its potential correspondence to the aims approach, as the central research question of the project.
+*[also state this earlier]*  Operationalized using science-health bigrams, this difference might show up in differences in noun frequency within or between sides (for example, opponents might use science bigrams more than supporters) or in different adjectives used with the same noun (for example, supporters might use different adjectives to describe health).
+
+[A1] had to set the project aside for a few years, but in January 2024 brought in the other authors.
+Without informing the other authors of the research question, all authors independently coded each comment in the sample as supporting or opposing ST, or too ambiguous to classify.
+Inter-rater reliability was good after one round of independent coding, and excellent after a round of discussion and separate review of discordant codes.
+
+Manual coding was limited to about 800 out of 22,000 comments.
+We used a machine learning model to scale up the manual codes (supporting/opposing), using them as training data and imputing or predicting codes for the remaining comments.
+To account for ambiguous comments (neither clearly supporting nor clearly opposing) and the possibility of error by the model, we conducted all analyses in parallel across three "codings": the *manual*ly-coded comments, the *imputed* codings for all 22k comments, and a *filtered* subset of the imputed codings, those where the model is at least 80% confident in its coding assignment.
+
+As we discuss below, it turned out that supporting comments were much less likely to use science-health bigrams than opposing comments, and so the bigram analysis omitted the majority of supporting comments.
+We decided to cover the whole corpus using a keyword search.
+But keyword searches are vulnerable to synonyms and actor-analyst conceptual mismatch.
+We therefore used word embeddings, specific to this corpus, to identify potential synonyms and check conceptual alignment; see @sec-embeddings.
+
+# Results {#sec-results}
 
@@
 Three bigrams occur in the majority of opposing comments, across all codings: "public health," "available science" (short for "best available science"), and "good science." Other common bigrams used by opponents include "human health," "environmental health," "reliable science," and "medical science."
+
+*[set blue L to like 40]*
 
@@
 
+*[already using red/blue for health/science]*
+
 ![15 most distinctive terms (unigrams) for supporters and opponents, by coding.
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 

### 2025-03-02 scripts/09_bigram_analysis.R

@@ -297,7 +297,7 @@ llr_plot = function(llr_df, title, n = 15) {
         ggplot(aes(bigram, ymax = llr, color = noun)) +
         geom_linerange(ymin = 0, linewidth = 1) +
         coord_flip() +
-        facet_grid(vars(side), ncol = 1, scales = 'free_y') +
+        facet_grid(vars(side), scales = 'free_y') +
         scale_color_brewer(palette = 'Set1') +
         labs(x = '', y = 'log likelihood ratio', title = title)
 }
@@ -310,6 +310,15 @@ p_df |>
                     caption = 'log(x + 10^-5)')
 ggsave(here(out_dir, '09_llr_top.png'), 
        height = 7, width = 8, bg = 'white', scale = 1.25)
+
+p_df$filtered |> 
+    llr() |> 
+    llr_plot('', n = 10) +
+    labs(title = 'Top bigrams, by log likelihood ratio', 
+            caption = 'log(x + 10^-5)')
+
+ggsave(here(out_dir, '09_llr_top_filtered.png'), 
+       height =  4, width = 6, bg = 'white')
 
 # opts = options(pillar.sigfig = 10)
 ## For manual, medians are identical bc the median bigram appears in 1 opposing doc and no supporting docs

### 2025-03-02 scripts/11_keyword_occurrence.R

@@ -14,6 +14,25 @@ source(here('R', 'extract_noun.R'))
 source(here('R', 'tab_out.R'))
 out_dir = here('out')
 
+load_text = function(coded_df) {
+    open_dataset(here('data', '03_text.parquet')) |>
+        inner_join(coded_df, by = 'comment_id') |> 
+        ## <https://github.com/apache/arrow/issues/40220>
+        collect()
+}
+
+count_tokens = function(coded_df) {
+    open_dataset(here('data', '03_annotations')) |>
+        inner_join(coded_df, by = 'comment_id') |> 
+        # filter(comment_id %in% coded_df$comment_id) |> 
+        group_by(comment_id) |> 
+        summarize(length = n()) |> 
+        inner_join(coded_df, by = 'comment_id') |> 
+        collect()
+}
+# count_tokens(coded$imputed)
+
+
 ## Load data ----
 data_dir = here('data')
 source(here('R', 'load_coding.R'))
@@ -22,12 +41,28 @@ coded = list(manual = load_manual_coding(filter_na = TRUE),
              filtered = load_imputed_fltd(), 
              imputed = load_imputed_full())
 
+## Doc length by support/opposition ----
+map(coded, count_tokens) |> 
+    bind_rows(.id = 'coding') |> 
+    ggplot(aes(length, group = support, color = support)) +
+    geom_freqpoly(bins = 80) +
+    scale_x_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
+    # guides(x = guide_axis_logticks()) +
+    annotation_logticks(sides = 'b') +
+    scale_color_brewer(palette = 'Set1') +
+    facet_wrap(vars(coding), 
+               ncol = 1, 
+               scales = 'free_y', 
+               axes = 'all_x')
+
+ggsave(here(out_dir, '11_len.png'), 
+       height = 3, width = 4, bg = 'white', scale = 2)
+
+
 ## Analyze keyword occurrence ----
 kword_occurence = function(coded_df) {
-    open_dataset(here('data', '03_text.parquet')) |>
-        inner_join(coded_df, by = 'comment_id') |> 
-        ## <https://github.com/apache/arrow/issues/40220>
-        collect() |>
+    coded_df |> 
+        load_text() |>
         mutate(science = str_detect(text, 
                                     regex('(?!(?-i)Regulatory Science)(scien)', 
                                           ignore_case = TRUE)), 

### 2025-04-28 paper/methods.qmd
starting to work on draft for IEEE; AFT format

@@
----
-format:
-  pdf:
-    mainfont: "Times New Roman"
-    fontsize: "12pt"
-    geometry: 
-      - "margin = 1in"
-    linestretch: 1.5
----
-
 ## Corpus assembly
@@
 For Strengthening Transparency, Regulations.gov reports 993,464 "comments received" and 22,390 "comments posted to this docket" and publicly available.
-*[TODO: email for clarification on relationship between these numbers]*  It's likely that this disparity is due, in part, to separately enumerating some mass comments that were submitted collectively.
-For example, comment 11822 (<https://www.regulations.gov/comment/EPA-HQ-OA-2018-0259-11822>) includes 22,786 individual comments collected by Friends of the Earth (likely using an online form) and submitted as large PDFs, one individual comment per page.
+As explained in the main text, this disparity is due to the way Regulations.gov counts identical mass comments separately for "comments received" but groups them together for "comments posted to this docket."  For example, comment 11822 (<https://www.regulations.gov/comment/EPA-HQ-OA-2018-0259-11822>) includes 22,786 individual comments collected by Friends of the Earth (likely using an online form) and submitted as large PDFs, one individual comment per page.
 
@@
 
-Comments were primarily submitted during two public comment periods, one following the publication of the Notice of Proposed Rulemaking (NPR) on *[date]* and the second in response to the Revised Notice of Proposed Rulemaking (RNPR) on *[date]*.
+Comments were primarily submitted during two public comment periods, one following the publication of the Notice of Proposed Rulemaking (NPR) on 2018-04-30 [@USEPAStrengtheningTransparencyRegulatory2018] and the second in response to the Revised Notice of Proposed Rulemaking (RNPR) on 202-03-03 [@USEPAStrengtheningTransparencyRegulatory2020a].
 Comments varied in length dramatically: a few handwritten postcards could not be parsed by OCR and were counted as being 0 words in length, and the longest comment (including the comment text and all attachment text) was 467,917 words long.
@@
 Next, the text was tokenized, stopwords removed *[source]*, and tokens lemmatized.
-The resulting token features were hashed to $1024 = 2^10^$ features, with signed binary values (0, +1, -1).
+The resulting token features were hashed to $1024 = 2^{10}$ features, with signed binary values (0, +1, -1).
 Zero variance features were removed.
@@
 This filtered subset contained 14,101 opposing comments (74%) and 4,876 supporting comments (26%), for a total of 18,977 (91% of the comments run through imputation).
+The distribution of document lengths is shown in @fig-len.
 
-*[The distribution of document lengths, in words (spaCy-annotated tokens), after coding is shown in fig @fig-len. 11_len.png]*
+![Distribution of document lengths, in words (spaCy-annotated tokens), by support/opposition and coding method](../out/11_len.png){#fig-len}
 

### 2025-04-28 paper/paper.qmd
starting to work on draft for IEEE; AFT format

@@
 authors:
-  - name: "D. J. Hicks"
+  - name: "D.J. Hicks"
     email: "dhicks4@ucmerced.edu"
     affiliations: 
-      - id: ucm
+      - id: a
         name: "University of California, Merced"
@@
     affiliations:
-      - ref: ucm
-    orcid: "*[todo]*"
+      - ref: a
+    orcid: "0000-0002-3066-2932"
   - name: "Cosmo Campbell"
     affiliations:
-      - ref: ucm
+      - ref: a
   - name: "Joseph Dad"
     affiliations:
-      - ref: ucm
+      - ref: a
 
@@
   pdf:
-    toc: true
+    toc: false
     number-sections: true
@@
     geometry: 
-      - "margin = 1in"
-    linestretch: 1.5
+      - "margin = 1.25in"
+    linestretch: 2
+  aft-pdf:
+    keep-tex: true
+    link-citations: true
+    mathspec: true
+    fig.env: "figure*"
+    include-in-header:
+        - text: |
+            \usepackage{pdflscape}
+    number-sections: true
+    cite-method: citeproc
 
@@
 
+In this paper, we argue that the aims approach to values in science can provide a useful framework for identifying certain kinds of deep disagreements in public scientific controversies.
+Specifically, by applying text mining methods to a large collection of public comments on an open data rule at the US Environmental Protection Agency (USEPA), we argue that supporters and opponents of the rule exemplify the two rival views of the aims of science identified by Hicks [-@HicksScientificPracticesTheir2012; -@HicksWhenVirtuesAre2022].
+
 *[
-- general science-and-values
-- aims approach
-- Strengthening Transparency
+- aims approach in general
+- Hicks' two views
 ]*
+
+## *Strengthening Transparency in Regulatory Science*
+
+*Strengthening Transparency in Regulatory Science* (henceforth Strengthening Transparency or ST) was a rule proposed^[Technically, the rule was finalized and adopted in January 2021, just a few weeks before Trump's first administration ended.
+The rule was never enforced, and formally vacated in February 2021 by a federal judge on procedural grounds and without objection from the Biden administration.
+The final title of the rule was *Strengthening Transparency in Pivotal Science Underlying Significant Regulatory Actions and Influential Scientific Information*.
+For simplicity, I usually refer to it by the original title, and I characterize its effects counterfactually.] by USEPA during the first Trump administration [@USEPAStrengtheningTransparencyRegulatory2018; @USEPAStrengtheningTransparencyPivotal2021].
+Strengthening Transparency would have imposed a strong open data and code requirement on any "regulatory science" utilized by USEPA to develop new regulations, generally prohibiting the agency from utilizing any research where the data and analysis code were not publicly available [@HicksWhenVirtuesAre2022; @HicksOpenScienceReplication2023].
+
+At least in the initial public notice, Strengthening Transparency was rationalized by an appeal to the open science movement and as a necessary response to the "replication crisis," a purportedly widespread crisis in which influential and high-profile scientific results often could not be replicated [@USEPAStrengtheningTransparencyRegulatory2018 18769-70].
+@HicksOpenScienceReplication2023 argued that, at the time ST was proposed, there was limited evidence that the replication crisis extended beyond a handful of scientific fields, and specifically there was no evidence that the crisis covered environmental epidemiology.
+
+Strengthening Transparency was the culmination of a campaign against "secret science" at USEPA.
+At least since the 1990s [@LernerRepublicansAreUsing2017], regulated industries have leveled complaints against certain highly influential environmental epidemiological studies [such as @DockeryAssociationAirPollution1993 and @RauhImpactPrenatalChlorpyrifos2006], alleging nonspecific data and analysis problems and demanding access to the "raw data" [@HicksWhenVirtuesAre2022 §2.2].
+*[secret science act]*
+
+Strengthening Transparency received a relatively high degree of mainstream news coverage [@HakimPesticideStudiesWon2018; @AchenbachScientistsDecryNew2018; @MufsonEPAExcludedTop2018; @FriedmanEPALimitScience2019; @DennisEPAPushesAhead2019] and was publicly opposed by a number of scientific organizations and environmental advocacy organizations *[cites]*.
+Notably, some of the public opponents included prominent advocates of open science *[cites]*.
+Perhaps the most common argument against ST, as reported in news coverage or made by public figures, was that environmental epidemiological studies are often based on medical records and other sensitive information that legally cannot be made publicly available.
+Another common concern was that, if the rule was applied retroactively, for studies conducted decades ago the "raw data" might no longer exist [again, such as @DockeryAssociationAirPollution1993 and @RauhImpactPrenatalChlorpyrifos2006].
+
+## Text mining
 
@@
 Broadly, text mining aims to identify patterns in the use of terms across a large collection of documents (called a *corpus*).
-While initially developed by computer scientists, text mining methods have been widely adopted by social scientists [@NelsonComputationalGroundedTheory2020; @StoltzMappingTextsComputational2024] and (more contentiously) humanists, under the umbrella of "digital humanities" [@UnderwoodDistantHorizonsDigital2019].
+While initially developed by computer scientists, text mining methods have been widely adopted by both social scientists [@NelsonComputationalGroundedTheory2020; @StoltzMappingTextsComputational2024] and — under the umbrella of "digital humanities" — humanists [@UnderwoodDistantHorizonsDigital2019].
 
@@
 Their most recent work combines machine translation with topic modeling, expanding the scope of their analysis to incorporate multiple languages.
+@HicksRaceScienceMainstream2024 use agnotology as a theoretical framework for a text mining analysis of race science in mainstream psychology.
 The text mining case studies in @DeBlockDynamicsScienceComputational2022 deploy a variety of methods, including named-entity recognition (NER) and network analysis [@PenceHowNotFight2022] and keyword search to facilitate deep reading [@GinammiBolzanoKantTraditional2022].
 @VaesenPotentialSupervisedMachine2022 argues that supervised machine learning approach can incorporate philosophical expertise in ways that topic modeling cannot; we independently adopted a similar approach in part of the current study.
-
-*[methods pieces by Charles?]*
 
@@
 
-We propose that public comments on high-profile governmental regulatory actions can be apt and convenient corpora for public scientific controversies.
+We propose that public comments on high-profile governmental regulatory actions such as Strengthening Transparency can be apt and convenient corpora for understanding the publics involved in public scientific controversies.
 In the United States, the Administrative Procedure Act requires federal government agencies to follow a "notice and comment" process before adopting new or modified regulations *[cite]*.
@@
 
-ST had an exceptionally high profile, with nearly one million comments.^[In the text mining results below, the corpus contains "only" about 22,000 comments.
+Most likely due to the mainstream news coverage, ST had an exceptionally high profile, with nearly one million comments submitted.^[In the text mining results below, the corpus contains "only" about 22,000 comments.
 Regulations.gov distinguishes between "comments posted" and "comments received"; a single comment for the "posted" count can include multiple comments for the "received" count.
@@
 This aggregation would count as one "posted" comment but tens of thousands of "received" comments.]  We use this substantial corpus to examine how supporters and opponents talk about science and health, finding that opponents talk about health and the environment about as often as they talk about science; while supporters talk exclusively about science, with a distinctive emphasis on a cynical or skeptical framing of the epistemic reliability of the science utilized by USEPA.
-We argue that these differences correspond to Hicks' [-@HicksWhenVirtuesAre2022] distinction between broad and narrow views of science.
 
-The bulk of this paper gives a non-technical overview of our text mining methods (@sec-methods-overview) and results (@sec-results).
-@sec-methods provides a more technical discussion of our methods, for readers who are familiar with text mining.
+We argue that these differences between opponents and supporters correspond to Hicks' [-@HicksWhenVirtuesAre2022] distinction between broad and narrow views of science.  *[more]*
+
+@sec-methods-overview gives a non-technical overview of our text mining methods; @sec-methods provides a more technical discussion, for readers who are familiar with text mining.
+@sec-results reports the empirical results of the analysis.
 In @sec-discussion, we first argue that our empirical results can be fruitfully understood in terms of the two views of the aims of science.
@@
 All agency-published documents and public comments associated with a proposed rule are collected together in a docket, e.g., EPA-HQ-OA-2018-0259 for Strengthening Transparency, available at <https://www.regulations.gov/docket/EPA-HQ-OA-2018-0259>.
-We collected all public comments from this docket in 2020 and 2021, with a final check several months after ST was finalized (in the last few weeks of the Trump administration) and vacated (a few weeks into the Biden administration).
+[A1] collected all public comments from this docket in 2020 and 2021, with a final check several months after ST was finalized (in the last few weeks of the Trump administration) and vacated (a few weeks into the Biden administration).
 
-Initial, undirected exploration of the corpus in 2020-2021 by [A1] indicated that (a) there was a qualitatively clear division between supporters and opponents of ST, (b) commenters offered substantive reasons for their position on ST, but (c) opposing comments were much, much more prevalent than supportive comments.
-The first point meant that manual qualitative coding could be used to identify supporting and opposing comments; combined with the second point, this meant that the corpus was suitable for contrasting the arguments given on either side of the ST controversy.
+Initial, undirected exploration of the corpus in 2020-2021 by [A1] indicated that (a) there was a qualitatively clear division between supporters and opponents of ST, (b) commenters offered substantive reasons for their position on ST, (c) these reasons often connected science to public health, but (d) opposing comments were much, much more prevalent than supportive comments.
+The point (a) meant that manual qualitative coding could be used to identify supporting and opposing comments; combined with points (b) and (c), this meant that the corpus was likely apt for contrasting the arguments given on either side of the ST controversy, and especially the way they connected (or not) epistemic and pragmatic aims.
 But the third point meant that a simple random sample would include very few supporters.
@@
 [A1] had also identified the relationship between science and health, and its potential correspondence to the aims approach, as the central research question of the project.
-*[also state this earlier]*  Operationalized using science-health bigrams, this difference might show up in differences in noun frequency within or between sides (for example, opponents might use science bigrams more than supporters) or in different adjectives used with the same noun (for example, supporters might use different adjectives to describe health).
+Operationalized using science-health bigrams, this difference might show up in differences in noun frequency within or between sides (for example, opponents might use science bigrams more than supporters) or in different adjectives used with the same noun (for example, supporters might use different adjectives to describe health).
 
@@
 read_rds(here(out_dir, '09_n_docs.Rds'))
-read_rds(here(out_dir, '09_n_docs_bigrams.Rds'))
+# read_rds(here(out_dir, '09_n_docs_bigrams.Rds'))
 ```
 
-In the manual coding, the 33 "neither" comments were almost all procedural requests to extend the public comment period, with a few comments discussing an issue that was brought up as an example in the first NPR (the linear no-threshold model for radiation hazard) that we did not consider relevant to our analysis.
+In the manual coding, the 33 "neither" comments were almost all procedural requests to extend the public comment period, with a few comments discussing an issue that was mentioned as an example in the first NPR (the linear no-threshold model for radiation hazard) that we did not consider relevant to our analysis.
 Because the ML classifier was not trained on these 33 "neither" comments, it forced all such comments into either the "support" or the "oppose" category.
@@
 A skeptical or cynical assessment of science would provide support for ST insofar as open data would make science less secretive and phony and more verifiable and true, etc.
-@HicksOpenScienceReplication2023 argued that there is no empirical evidence of a replication crisis in environmental epidemiology [though see @BagiletEstimatingSmallEffects], and that ST's open data requirements would not mitigate widespread irreplicability even if it were a problem in that field.
+@HicksOpenScienceReplication2023 argued that there is no empirical evidence of a replication crisis in environmental epidemiology [though see @BagiletEstimatingSmallEffects *[cite as well in intro]*], and that ST's open data requirements would not mitigate widespread irreplicability even if it were a problem in that field.
+@HicksDevelopingMeasuresPublic2024 report preliminary evidence that cynicism is a better predictor of generalized trust in science than the value-free ideal.
 
@@
 
+```{=latex}
+\begin{landscape}
+```
 ```{r}
@@
 #| tbl-cap: "Keyword occurrence, by coding method and support. Footnotes indicate regular expressions used for search."
+#| tbl-align: center
 read_rds(here(out_dir, '11_keyword_tab.Rds'))
+```
+```{=latex}
+\end{landscape}
 ```
@@
 
-We argue that the results above can be fruitfully understood in terms of *[Hicks']* account of two views of the aims of science.
+We argue that the results above can be fruitfully understood in terms of @HicksWhenVirtuesAre2022 account of two views of the aims of science.
 Specifically, we claim that supporters deployed the narrow view, while opponents deployed the broad view.
+
 On the narrow view of supporters of ST, science has only epistemic aims.
@@
 
-By contrast, opponents deployed the broad view, on which science has both epistemic and practical aims.
-Specifically, they appealed to the practical aims of protecting human health and the environment, raising concerns that ST's open data requirements would frustrate the pursuit of these aims, and was therefore "anti-science" *[Hicks]*.
+By contrast, opponents of ST deployed the broad view, on which science has both epistemic and practical aims.
+Specifically, they appealed to the practical aims of protecting human health and the environment, raising concerns that ST's open data requirements would frustrate the pursuit of these aims, and was therefore "anti-science" [a primary goal of @HicksWhenVirtuesAre2022 is to develop a robust version of this argument].
 In this way, the broad view provides an argumentative challenge to ST.
@@
 On Hicks' theory of value, protecting human health and the environment are aims of environmental public health, but the impacts of regulation on the economy are still accidents, and thus irrelevant to the standards of good scientific research.
-@HicksValuesDisclosuresTrust2022 provide survey evidence that, in the context of chemical regulation, members of the public put more trust in a scientist who discloses public health values than one who does not disclose values or discloses economic values.
+@HicksValuesDisclosuresTrust2022 provide survey evidence that, in the context of chemical regulation, members of the public — including political conservatives — put more trust in a scientist who discloses public health values than one who does not disclose values or discloses economic values.
 
@@
 
+# References
+
+::: {#refs}
+:::
+
 \appendix
+\clearpage
+\pagenumbering{arabic}
+\renewcommand*{\thepage}{S\arabic{page}}
+\renewcommand\thefigure{S\arabic{figure}}    
+\setcounter{figure}{0} 
+\renewcommand\thetable{S\arabic{table}}    
+\setcounter{table}{0}
+\renewcommand*{\thesection}{S\arabic{section}}
 

### 2025-05-07 paper/paper.qmd
working towards draft

@@
 
-*[
-- aims approach in general
-- Hicks' two views
-]*
+## The aims approach
+
+While not as influential as the argument from inductive risk *[Douglas]*, the aims approach to values in science provides a distinct challenge to the value-free ideal.
+Proponents of the approach argue that science typically has not only epistemic aims but also practical or pragmatic ones, and that at any stage of inquiry decisions should be made so as to promote those aims.
+So, specifically, in the "core" of inquiry decisions should be made so as to promote pragmatic aims, and in this way non-epistemic values can play a legitimate role in the "core" of inquiry.
+*[cites]*  For example, *[Elliott and McKaughan]* argue that the pragmatic aim of quick and protective regulatory review supports using *[high-throughput]* methods and a potentially greater false positive rate in chemical safety testing.
+Note that this argument does not necessarily deny that science has epistemic aims, as may be done by approaches inspired by John Dewey *[Anderson; Brown]* or Donald Davidson *[Clough]*.
+This argument is also independent of questions about the definition of epistemic values or a sharp demarcation of values into "epistemic" vs.
+"social" *[Longino; Rooney; McMullin; Steel]*.
+Instead, the key assumption of this argument is that science does not have only epistemic aims.
+
+Proponents of the aims approach often argue for pluralism about aims of science *[Elliott and McKaughan; Potochnik]*, with *[Lusk and Elliott]* explicitly rejecting "the view
+that science has global aims that constrain the influence of local aims on scientific
+practice" (1).
+Both epistemic and pragmatic aims are often presented as specific to a discipline, field, or research program *[Tapestry ch.
+4; examples]*.
+*[Intemann's]* early statement of the approach argues for "democratically endorsed" aims.
+*[more?]*
+
+A distinctive aspect of *[Hicks']* version of the aims approach is an analysis of two general "views" of the aims of science, and a claim that an unrecognized disagreement between the two views is a recurrent feature of debates, not only among scientists and the public but also among philosophers.
+For Hicks, the "broad view" corresponds to the aims approach argument sketched above — science has both epistemic and pragmatic aims — while the "narrow view" denies the "key assumption" of that argument, taking science to have only epistemic aims *[cites]*.
+On the narrow view, applications of scientific research are seen as happy accidents but not the primary function or purpose of research.
+Allowing pragmatic considerations (read: non-epistemic values) to influence inquiry is a corrupting influence, morally on a par with the epistemic distortions introduced by industrial funding *[Bennett and Holman; Fernandez Pinto]* or *[Wansink]*.
+
+Something like Hicks' two views appears in philosophical debates over the argument from inductive risk. *[acceptance vs. belief]*
 
@@
 
-An inductive risk framing might work well for opponents, who are very likely to talk about public health, the environment, and ecology.
-But it does not seem to fit supporters, who talk about the economy about as often as they talk about the environment, and are less likely than not to talk about regulation.
+An inductive risk framing might work well for opponents, who are very likely to talk about public health, the environment, ecology, and other downstream social implications of ST.
+But it does not seem to fit supporters, who talk about the economy about as often as they talk about the environment — that is, very little — and are less likely than not to talk about regulation.
 We therefore take the aims approach to be a better fit for analyzing this particular controversy.
@@
 ## The views as interpretive frames, not beliefs
+
+When we use text mining methods to study public comments like this, what are we studying?
+In particular, when we attribute one view of the aim of science to members of the public, are we saying that they *believe* this view, in anything like the way philosophers typically use that term?
+
+First, consider the many "form letter" comments.
+We labeled these comments in this way because we believe they were generated using an online form, which provided editable letter text and fields for a prospective commentator to enter their name, city and state, etc.
+These comments comprise 98% of the comments on Strengthening Transparency (counted granularly: 1 - 22k / 1M).
+While the text of these letters was editable, almost all prospective commentators made no changes.
+It is plausible that the vast majority of form letter commentators did not carefully read the letter they were "writing."
+
+Second, belief is often understood by philosophers as either a felt commitment to the truth of a claim or proposition, or as a behavioral disposition (with certain cognitive aspects) *[Schwitzgebel]*.
+In either conceptualization, beliefs are expected to be more-or-less stable across contexts (keeping fixed available evidence, belief in logically relevant claims, etc.).
+But there is circumstantial evidence^[By this we mean that the evidence below does not show that any identifiable individual endorsed different views of the aims of science in two different contexts.
+Instead our argument is that two incompatible views appear to be widely endorsed in the same group in different contexts, and thus it is highly probably that many individual members of this group have or would endorse both views in an unstable or logically inconsistent way.] that the views of aims of science we identified here are not stable.
+
+Consider opponents of ST, who we take to generally be environmentalists and political liberals or progressives ("liberals" for short).
+While liberals utilized the broad view in the ST controversy, in 2021 liberals also supported the development of scientific integrity plans across the Biden administration *[cites]*.
+The Biden administration's framework for these plans utilized the narrow view and the value-free ideal, portraying non-epistemic considerations exclusively as political interference or inappropriate influence (and without defining either of these terms):
+
+> Scientific findings and products must not be suppressed, delayed, or altered for political purposes and must not be subjected to inappropriate influence. 
+> 
+> Science, and public trust in science, thrives in an environment that shields scientific data and analyses and their use in policymaking from political interference or inappropriate influence.
+[@OSTPReleasesFramework2023 29-30]
+
+On the other hand, as discussed above, we considered the possibility that supporters of ST — who we take to generally be political conservatives — appealed to pragmatic aims such as promoting economic growth, supporting small businesses, or reducing excessive regulation on industry.
+This would have amounted to utilizing the broad view, just with a different set of pragmatic aims from the liberal version.
+We found no evidence of a conservative broad view in the ST controversy; but of course conservatives appeal to such pragmatic aims in many other contexts *[cites]*.
+
+For these reasons, we do not interpret the comments on ST as expressing beliefs of the public on the aims of science.
+As an alternative, we suggest that the narrow and broad views can be understood as interpretive frames [@KleinDataFrameTheory2007].
+Based on studies of the ways professionals make decisions in complex situations (primarily military intelligence analysts), @KleinDataFrameTheory2007 proposed that decisionmakers engage in "sensemaking," a form of inquiry [compare @BrownScienceMoralImagination2020 ch.
+1] aimed at understanding rather than empirical accuracy [@KleinDataFrameTheory2007], and specifically understanding to support effective decisionmaking in the face of practical uncertainty.
+A "frame" is any "explanatory structure[] that account[s] for the data," including but not limited to stories, scripts, and maps [@KleinDataFrameTheory2007 120].
+@KleinDataFrameTheory2007 emphasize that there is an iterative relationship between data and frame: "The data identify the relevant frame, and the frame determines which data are noticed.
+Neither of these comes first" [118; compare *[Hanson, Chang]*].
+Because individuals operate with different "repertoires of frames," and data-gathering is informed by frames, "Different people viewing the same events can perceive and recall different things depending on their goals and experiences" [@KleinDataFrameTheory2007 121].
+Extending this thought, we suggest that an individual's repertoire might include incompatible frames, and so similar events might be perceived and interpreted in very different ways by the same individual.
+
+Applied to the ST controversy, we suggest that both liberals and conservatives have both narrow and broad views of the aims of science among their repertoire of frames for interpreting public scientific controversies.
+*[narrow view is depoliticizing; broad view is politicizing?]*
+
+Reading the public comments as expressions of interpretive frames rather than beliefs can also help us understand what the "authors" of form letter comments were doing when they submitted a comment without modifying or even carefully reading the suggested text.
+For the low-stakes action of submitting a comment, details of argument and wording were less important than the general sentiment of the text — its vibes.
+So a few seconds to skim the suggested text would be sufficient for most commentators.
 

### 2025-05-09 paper/paper.qmd
enough content is done for ISEE; need to work on refs

@@
     number-sections: true
-  pdf:
-    toc: false
-    number-sections: true
-    link-citations: true
-    pdf-engine: "lualatex"
-    mainfont: "Times New Roman"
-    fontsize: "12pt"
-    geometry: 
-      - "margin = 1.25in"
-    linestretch: 2
   aft-pdf:
@@
             \usepackage{pdflscape}
+            \usepackage{framed}
     number-sections: true
@@
   echo: false
-  warning: true
+  warning: false
   message: false
@@
 
-While not as influential as the argument from inductive risk *[Douglas]*, the aims approach to values in science provides a distinct challenge to the value-free ideal.
+While not as influential as the argument from inductive risk *[Douglas, Havstad]*, the aims approach to values in science provides a distinct challenge to the value-free ideal.
 Proponents of the approach argue that science typically has not only epistemic aims but also practical or pragmatic ones, and that at any stage of inquiry decisions should be made so as to promote those aims.
 So, specifically, in the "core" of inquiry decisions should be made so as to promote pragmatic aims, and in this way non-epistemic values can play a legitimate role in the "core" of inquiry.
-*[cites]*  For example, *[Elliott and McKaughan]* argue that the pragmatic aim of quick and protective regulatory review supports using *[high-throughput]* methods and a potentially greater false positive rate in chemical safety testing.
+*[cites]*  For example, *[Elliott and McKaughan]* argue that the pragmatic aim of quick and protective regulatory review in chemical safety testing supports using rapid screening techniques, which are much faster than other methods but will likely have a higher false positive rate.
+
 Note that this argument does not necessarily deny that science has epistemic aims, as may be done by approaches inspired by John Dewey *[Anderson; Brown]* or Donald Davidson *[Clough]*.
@@
 "social" *[Longino; Rooney; McMullin; Steel]*.
-Instead, the key assumption of this argument is that science does not have only epistemic aims.
+Instead, the key assumption of this argument is that science does not have *only* epistemic aims.
 
@@
 
-At least in the initial public notice, Strengthening Transparency was rationalized by an appeal to the open science movement and as a necessary response to the "replication crisis," a purportedly widespread crisis in which influential and high-profile scientific results often could not be replicated [@USEPAStrengtheningTransparencyRegulatory2018 18769-70].
+At least in the initial public notice, proponents of Strengthening Transparency appealed to the open science movement and, specifically, open science as a necessary response to the "replication crisis," a purportedly widespread crisis in which influential and high-profile scientific results often could not be replicated [@USEPAStrengtheningTransparencyRegulatory2018 18769-70].
 @HicksOpenScienceReplication2023 argued that, at the time ST was proposed, there was limited evidence that the replication crisis extended beyond a handful of scientific fields, and specifically there was no evidence that the crisis covered environmental epidemiology.
@@
 
-Text mining is an emerging method for empirical philosophy of science.
+Text mining has emerged as a method for empirical philosophy of science over the past decade.
 Broadly, text mining aims to identify patterns in the use of terms across a large collection of documents (called a *corpus*).
@@
 Agencies are required to publish a Notice of Proposed Rulemaking (NPR), including the draft text of the rule and background information to justify it, and then collect and respond to public comments before finalizing the rule.
-Conveniently for our purposes, most comments are made freely and publicly available.
+Conveniently, most comments are made freely and publicly available on the website `regulations.gov`, which also includes a free API (application programming interface) to facilitate bulk downloads ("scraping").
 High-profile proposed rules can collect large numbers of comments, from a variety of different groups and perspectives on the issue at hand in the rule, making it apt for understanding how different groups might, for instance, deploy different views of the aims of science.
 
-Most likely due to the mainstream news coverage, ST had an exceptionally high profile, with nearly one million comments submitted.^[In the text mining results below, the corpus contains "only" about 22,000 comments.
+## Text mining ST: Primary findings and roadmap of the paper
+
+Presumably due to the mainstream news coverage, ST had an exceptionally high profile, with nearly one million comments submitted and available on `regulations.gov`.^[In the text mining results below, the corpus contains "only" about 22,000 comments.
 Regulations.gov distinguishes between "comments posted" and "comments received"; a single comment for the "posted" count can include multiple comments for the "received" count.
@@
 
-We argue that these differences between opponents and supporters correspond to Hicks' [-@HicksWhenVirtuesAre2022] distinction between broad and narrow views of science.  *[more]*
+We argue that these differences between opponents and supporters correspond to Hicks' [-@HicksWhenVirtuesAre2022] distinction between broad and narrow views of science.
 
@@
 In @sec-discussion, we first argue that our empirical results can be fruitfully understood in terms of the two views of the aims of science.
-We go on to suggest that, in the context of public discourse, deployment of philosophical ideas should be understood as somewhat opportunistic interpretive frames, rather than authentic expressions of beliefs or deep philosophical commitments.
-Finally, we consider the role of advocacy organizations in shaping the content of public scientific controversies, and consider the applicability of a "participatory" model developed in misinformation studies to the case of ST.
+We go on to suggest that, in the context of public discourse, deployment of philosophical ideas — such as views of the aims of science — might best be understood as somewhat opportunistic interpretive frames, rather than expressions of beliefs.
 
@@
 
-The US General Services Administration receives and publishes public comments on proposed rules using the website Regulations.gov (<https://www.regulations.gov>).
+The US General Services Administration receives and publishes public comments on proposed rules using the website `regulations.gov` (<https://www.regulations.gov>).
 Commenters may submit comments by online form, email, or physical mail, and electronic formats allow commenters to include a number of attachments.
-Attachments are typically detailed comments or to include journal articles, reports, etc., as supporting evidence.
+Attachments are typically longer comments or journal articles, reports, etc., as supporting evidence.
 
@@
 Initial, undirected exploration of the corpus in 2020-2021 by [A1] indicated that (a) there was a qualitatively clear division between supporters and opponents of ST, (b) commenters offered substantive reasons for their position on ST, (c) these reasons often connected science to public health, but (d) opposing comments were much, much more prevalent than supportive comments.
-The point (a) meant that manual qualitative coding could be used to identify supporting and opposing comments; combined with points (b) and (c), this meant that the corpus was likely apt for contrasting the arguments given on either side of the ST controversy, and especially the way they connected (or not) epistemic and pragmatic aims.
-But the third point meant that a simple random sample would include very few supporters.
+Example comments are given in figures @fig-opposed and @fig-support.
+
+::: {#fig-opposed}
+\begin{framed}
+\raggedright
+
+Im writing to speak out AGAINST your proposed regulations to assault the EPAs use of science in the middle of a global public health crisis.
+
+Peer-reviewed studies assessing the effects of pollution exposure on humans, based on actual personal health data which by law must be kept confidential are among the best science available to the Agency.
+The misguided proposal would direct EPA to ignore these data unless those confidential personal health data were made publicly available.
+
+This is both unnecessary and impractical. As a retired scientist, I urge you to abandon this reckless plan, for the sake of our childrens health. To do otherwise is unconscionable.
+\end{framed}
+
+Example of an opposing comment on Strengthening Transparency. Original available at <https://www.regulations.gov/comment/EPA-HQ-OA-2018-0259-19410>. 
+:::
+::: {#fig-support}
+\begin{framed}
+\raggedright
+I sincerely hope that you can distance the EPA agency from politically correct science. The only way to do this is to adopt the following methodology.
+
+The scientific method demands that researchers' work be fully transparent.
+
+Data must be made available for examination.
+
+Theories must be open to challenge.
+
+Disproven theories must be altered or set aside.
+
+Barriers must be removed that would prevent valid conclusions from rising to the top.
+
+I for one am tired of hearing about how bad technologies are for the planet, without any scientific proof that things today are different.
+Don't tell me that glarciers are melting and the ocean is rising a foot every 10 years.
+Remove yourselves from such statements and give facts about such a change: volume of water, how will it be measured, what could be causing it, what can be done that is both environmentally and economically sound.
+\end{framed}
+
+Example of a supporting comment on Strengthening Transparency. Original available at <https://www.regulations.gov/comment/EPA-HQ-OA-2018-0259-12305>.
+:::
+
+Point (a) meant that manual qualitative coding could be used to identify supporting and opposing comments; combined with points (b) and (c), this meant that the corpus was likely apt for contrasting the arguments given on either side of the ST controversy, and especially the way they connected (or not) epistemic and pragmatic aims.
+But (d) meant that a simple random sample would include very few supporters, which would be suboptimal for using machine learning to automatically code the entire corpus.
 @sec-sampling details the sampling procedure that we used for the manual coding step.
@@
 #|   - "Comments with science/health bigrams"
-#| layout-cols: 2
+#| layout-ncol: 1
 read_rds(here(out_dir, '09_n_docs.Rds'))
-# read_rds(here(out_dir, '09_n_docs_bigrams.Rds'))
+read_rds(here(out_dir, '09_n_docs_bigrams.Rds'))
 ```
@@
 
+*[move most of this to discussion]*
 There are also notable differences in the overall tone of the adjectives in supporters' and opponents' top bigrams.
@@
 Deployment of the broad view might also explain why opponents said relatively little about business and the economy.
-On Hicks' theory of value, protecting human health and the environment are aims of environmental public health, but the impacts of regulation on the economy are still accidents, and thus irrelevant to the standards of good scientific research.
+On Hicks' analysis protecting human health and the environment are aims of environmental public health, but the impacts of regulation on the economy are still accidents, and thus irrelevant to the standards of good scientific research.
 @HicksValuesDisclosuresTrust2022 provide survey evidence that, in the context of chemical regulation, members of the public — including political conservatives — put more trust in a scientist who discloses public health values than one who does not disclose values or discloses economic values.
@@
 Applied to the ST controversy, we suggest that both liberals and conservatives have both narrow and broad views of the aims of science among their repertoire of frames for interpreting public scientific controversies.
-*[narrow view is depoliticizing; broad view is politicizing?]*
+But the frames were evoked differently in this particular controversy by being connected to two very different frames on the general trustworthiness of EPA/regulatory science *[McCright and Dunlap]*
 
-Reading the public comments as expressions of interpretive frames rather than beliefs can also help us understand what the "authors" of form letter comments were doing when they submitted a comment without modifying or even carefully reading the suggested text.
+While the degree of conservative distrust in science is often exaggerated *[cites]*, there is a large difference in partisan attitudes towards USEPA: @CerdaAmericansFeelFavorably2024 reports that only 32% of Republican/Lean Republican survey respondents have a favorable attitude towards USEPA, compared to 73% of Democratic/Lean Democratic respondents.
+Republican distrust of USEPA appears in the bigram analysis *[xref]*, in the *way* supporters talk about science, using adjectives such as *[list]*.
+These characterizations of science portray it as suffering from an inappropriate influence of non-epistemic values — specifically, politicization *[cite]*.
+The narrow view and the value-free ideal justify outrage at this situation, and by extension policies such as ST that are (believed to) mitigate these inappropriate influences.
+We further suggest that liberal utilization of the narrow view, in the context of scientific integrity plans, is also a response to politicization.
+
+More generally, the narrow view might be evoked or deployed when science is (perceived as) being influenced by *values one disagrees with*.
+This can be understood as what we might call a *depoliticization rhetorical strategy*, in two senses: first, as a call to mitigate the influence of politics; but second as an appeal to premises (epistemic values) that are taken to be uncontroversial and external to politics.
+In the second sense, depoliticization is a version of proxy politics [@HicksScientificControversiesProxy2017] or what Sarewitz calls the "scientization" of politics [@SarewitzHowScienceMakes2004], where political actors present political concerns and arguments as apolitical and purely scientific.
+
+*[broad view]*
+
+Finally, reading the public comments as expressions of interpretive frames rather than beliefs can also help us understand what the "authors" of form letter comments were doing when they submitted a comment without modifying or even carefully reading the suggested text.
 For the low-stakes action of submitting a comment, details of argument and wording were less important than the general sentiment of the text — its vibes.
 So a few seconds to skim the suggested text would be sufficient for most commentators.
-
-## The role of advocacy organizations
-
-*[Starbird: participatory disinformation]*
 

### 2025-05-10 paper/paper.qmd
bibfile

@@
 acknowledgments: "acknowledgments"
+
+bibliography: "st_text_mining.yaml"
 
@@
 <!-- 
+PHOS submission guidelines: 
+<https://www.cambridge.org/core/journals/philosophy-of-science/information/author-instructions>
+
+Articles are limited to 9000 words, inclusive of abstract, footnotes, in-text citations, and appendices intended to appear with the manuscript.
+Reference sections and online-only appendices are not included in the word count.
+
 BJPS submission guidelines: 

### 2025-05-11 paper/paper.qmd
working on refs

@@
 
-While not as influential as the argument from inductive risk *[Douglas, Havstad]*, the aims approach to values in science provides a distinct challenge to the value-free ideal.
+While not as influential as the argument from inductive risk [@DouglasInductiveRiskValues2000; @HavstadSensationalScienceArchaic2021], the aims approach to values in science provides a distinct challenge to the value-free ideal.
 Proponents of the approach argue that science typically has not only epistemic aims but also practical or pragmatic ones, and that at any stage of inquiry decisions should be made so as to promote those aims.
 So, specifically, in the "core" of inquiry decisions should be made so as to promote pragmatic aims, and in this way non-epistemic values can play a legitimate role in the "core" of inquiry.
-*[cites]*  For example, *[Elliott and McKaughan]* argue that the pragmatic aim of quick and protective regulatory review in chemical safety testing supports using rapid screening techniques, which are much faster than other methods but will likely have a higher false positive rate.
+*[cites]*  For example, @ElliottNonepistemicValuesMultiple2014 argue that the pragmatic aim of quick and protective regulatory review in chemical safety testing supports using rapid screening techniques, which are much faster than other methods but will likely have a higher false positive rate.
 
-Note that this argument does not necessarily deny that science has epistemic aims, as may be done by approaches inspired by John Dewey *[Anderson; Brown]* or Donald Davidson *[Clough]*.
+Note that this argument does not necessarily deny that science has epistemic aims, as may be done by approaches inspired by John Dewey [@AndersonUsesValueJudgments2004; @BrownScienceMoralImagination2020] or Donald Davidson [@CloughRacistValueJudgments2008; @CloughUsingValuesEvidence2020].
 This argument is also independent of questions about the definition of epistemic values or a sharp demarcation of values into "epistemic" vs.
-"social" *[Longino; Rooney; McMullin; Steel]*.
+"social" [@McMullinValuesScience1982; @RooneyValuesScienceEpistemic1992; @LonginoGenderPoliticsTheoretical1995; @SteelEpistemicValuesArgument2010].
 Instead, the key assumption of this argument is that science does not have *only* epistemic aims.
 
-Proponents of the aims approach often argue for pluralism about aims of science *[Elliott and McKaughan; Potochnik]*, with *[Lusk and Elliott]* explicitly rejecting "the view
-that science has global aims that constrain the influence of local aims on scientific
-practice" (1).
-Both epistemic and pragmatic aims are often presented as specific to a discipline, field, or research program *[Tapestry ch.
+Proponents of the aims approach often argue for pluralism about aims of science [@ElliottNonepistemicValuesMultiple2014; @PotochnikDiverseAimsScience2015; @PotochnikIdealizationAimsScience2017], with @LuskGlobalAimsScience2024 explicitly rejecting "the view that science has global aims that constrain the influence of local aims on scientific practice" (1).
+Both epistemic and pragmatic aims are often presented as specific to a discipline, field, or research program *[@ElliottTapestryValuesIntroduction2017 ch.
 4; examples]*.
-*[Intemann's]* early statement of the approach argues for "democratically endorsed" aims.
+@IntemannDistinguishingLegitimateIllegitimate2015 coined the term "aims approach," and argued for "democratically endorsed epistemological and social aims of research" (2), which will be pluralistic insofar as different polities democratically endorse different aims.
 *[more?]*
 
-A distinctive aspect of *[Hicks']* version of the aims approach is an analysis of two general "views" of the aims of science, and a claim that an unrecognized disagreement between the two views is a recurrent feature of debates, not only among scientists and the public but also among philosophers.
+A distinctive aspect of @HicksWhenVirtuesAre2022 version of the aims approach is an analysis of two general "views" of the aims of science, and a claim that an unrecognized disagreement between the two views is a recurrent feature of debates, not only among scientists and the public but also among philosophers.
 For Hicks, the "broad view" corresponds to the aims approach argument sketched above — science has both epistemic and pragmatic aims — while the "narrow view" denies the "key assumption" of that argument, taking science to have only epistemic aims *[cites]*.
 On the narrow view, applications of scientific research are seen as happy accidents but not the primary function or purpose of research.
-Allowing pragmatic considerations (read: non-epistemic values) to influence inquiry is a corrupting influence, morally on a par with the epistemic distortions introduced by industrial funding *[Bennett and Holman; Fernandez Pinto]* or *[Wansink]*.
+Allowing pragmatic considerations (read: non-epistemic values) to influence inquiry is a corrupting influence and a form of research misconduct, *morally* on a par with the epistemic distortions introduced by industrial funding [@HolmanExperimentationIndustrialSelection2017; @FernandezPintoEpistemicDiversityIndustrial2023] or fraud [@BrightWhyScientistsLie2021].
 
@@
 
-*[move most of this to discussion]*
-There are also notable differences in the overall tone of the adjectives in supporters' and opponents' top bigrams.
-Supporters seem to express skepticism or cynicism about the quality of scientific research: science is described as secret, secretive, phony, false, fake, shady, or deceptive.
-In the context of these skeptical assessments, seemingly-positive valence bigrams — such as real, true, verifiable, and sound science — might be used as privatives, that is, expressing the idea that actually-existing science is *not* real, true, verifiable, or sound.
-A skeptical or cynical assessment of science would provide support for ST insofar as open data would make science less secretive and phony and more verifiable and true, etc.
-@HicksOpenScienceReplication2023 argued that there is no empirical evidence of a replication crisis in environmental epidemiology [though see @BagiletEstimatingSmallEffects *[cite as well in intro]*], and that ST's open data requirements would not mitigate widespread irreplicability even if it were a problem in that field.
-@HicksDevelopingMeasuresPublic2024 report preliminary evidence that cynicism is a better predictor of generalized trust in science than the value-free ideal.
-
 Table @tbl-noun-occ and figure @fig-noun-occ examine the occurrence of bigrams, aggregated by noun, for supporters and opponents.
@@
 As an alternative to the aims approach, our findings might be interpreted in terms of inductive risk, and specifically differences between supporters and opponents in the relative importance of over- vs.
-under-regulation *[cites]*.
+under-regulation [compare @HicksScientificControversiesProxy2017 on climate change].
 That is, supporters of ST might put more weight on over-regulation and preventable economic harms; while opponents of ST might put more weight on under-regulation and preventable disease and loss of life due to pollution.
@@
 
-Second, belief is often understood by philosophers as either a felt commitment to the truth of a claim or proposition, or as a behavioral disposition (with certain cognitive aspects) *[Schwitzgebel]*.
+Second, belief is often understood by philosophers as either a felt commitment to the truth of a claim or proposition, or as a behavioral disposition (with certain cognitive aspects) [@SchwitzgebelBelief2024].
 In either conceptualization, beliefs are expected to be more-or-less stable across contexts (keeping fixed available evidence, belief in logically relevant claims, etc.).
@@
 @KleinDataFrameTheory2007 emphasize that there is an iterative relationship between data and frame: "The data identify the relevant frame, and the frame determines which data are noticed.
-Neither of these comes first" [118; compare *[Hanson, Chang]*].
+Neither of these comes first" [118\; compare @HansonPatternsDiscoveryInquiry1958; @ChangInventingTemperatureMeasurement2004].
 Because individuals operate with different "repertoires of frames," and data-gathering is informed by frames, "Different people viewing the same events can perceive and recall different things depending on their goals and experiences" [@KleinDataFrameTheory2007 121].
@@
 Applied to the ST controversy, we suggest that both liberals and conservatives have both narrow and broad views of the aims of science among their repertoire of frames for interpreting public scientific controversies.
-But the frames were evoked differently in this particular controversy by being connected to two very different frames on the general trustworthiness of EPA/regulatory science *[McCright and Dunlap]*
+But the frames were evoked differently in this particular controversy by being connected to two very different frames on the general trustworthiness of EPA/regulatory science [@McCrightAntireflexivity2010; @McCrightPoliticalPolarizationSupport2014].
 
 While the degree of conservative distrust in science is often exaggerated *[cites]*, there is a large difference in partisan attitudes towards USEPA: @CerdaAmericansFeelFavorably2024 reports that only 32% of Republican/Lean Republican survey respondents have a favorable attitude towards USEPA, compared to 73% of Democratic/Lean Democratic respondents.
-Republican distrust of USEPA appears in the bigram analysis *[xref]*, in the *way* supporters talk about science, using adjectives such as *[list]*.
-These characterizations of science portray it as suffering from an inappropriate influence of non-epistemic values — specifically, politicization *[cite]*.
+Republican distrust of USEPA appears in the bigram analysis *[xref]*, in the *way* supporters talk about science.
+Namely, supporters seem to express skepticism or cynicism about the quality of scientific research: science is described as secret, secretive, phony, false, fake, shady, or deceptive.
+In the context of these skeptical assessments, seemingly-positive valence bigrams — such as real, true, verifiable, and sound science — might be used as privatives, that is, expressing the idea that actually-existing science is *not* real, true, verifiable, or sound.
+A skeptical or cynical assessment of science would provide support for ST insofar as open data would make science less secretive and phony and more verifiable and true, etc.^[@HicksOpenScienceReplication2023 argued that there is no empirical evidence of a replication crisis in environmental epidemiology [though see @BagiletEstimatingSmallEffects *[cite as well in intro]*], and that ST's open data requirements would not mitigate widespread irreplicability even if it were a problem in that field.] @HicksDevelopingMeasuresPublic2024 report preliminary evidence that cynicism is a better predictor of generalized trust in science than the value-free ideal.
+
+This cynical characterization of science portrays it as suffering from an inappropriate influence of non-epistemic values — specifically, politicization [@BrownScienceDemocracyExpertise2009 ch.
+8; @DouglasDifferentiatingScientificInquiry2023].
 The narrow view and the value-free ideal justify outrage at this situation, and by extension policies such as ST that are (believed to) mitigate these inappropriate influences.

### 2025-06-04 talk/talk.qmd
talk files

---
title: "The aims of science and an open science controversy at US EPA"
authors: "**D. J. Hicks** (they/them), Emilio Lobato, Cosmo Campbell, and Joseph Dad"
institute: |
  | University of California, Merced
  | ![](qr_code.png)
  | [tinyurl.com/25y8hn8h](https://tinyurl.com/25y8hn8h)

format: 
    revealjs:
        width: 1.1*1440
        height: 1080
        toc: true
        toc-depth: 1
        toc-title: Outline
        slide-number: true
        self-contained: true
        theme: [default, custom.scss]
        center: false
        footnotes-hover: true
        reference-location: document
        include-in-header: append-title-slide.html
        auto-stretch: false
link-citations: true
bibliography: ['/Users/danhicks/Google Drive/Writing/viss/paper/VISS.yaml']
---

# *Strengthening Transparency in Regulatory Science*

## *Strengthening Transparency in Regulatory Science*

- Strong open science rule at USEPA
- First Trump administration
- Proposed in 2018, briefly adopted in January 2021
- Based on Secret Science Act, industry complaints about public health studies

## Public comments on ST

- Notice and Comment process
    - Required under Administrative Procedure Act (1946)
    - Submitted by mail, email, or web form
    - Almost always made available on [regulations.gov](https://www.regulations.gov/) →\
        Easy to retrieve for text mining

- Public comments on ST
    - ~1M counting very granularly
    - ~21k as grouped on regulations.gov
    - Opponents generally political liberals
    - Supporters generally political conservatives

## Example comment: Opponent

Im writing to speak out AGAINST your proposed regulations to assault the EPAs use of [science]{.blue} in the middle of [a global public health crisis]{.red}.</br></br>

[Peer-reviewed studies]{.blue} assessing the effects of pollution exposure on humans, based on actual personal [health]{.red} [data]{.blue} which by law must be kept confidential are among [the best science available]{.blue} to the Agency. The misguided proposal would direct EPA to [ignore these data]{.blue} unless those confidential personal [health]{.red} [data]{.blue} were made publicly available.</br></br>

This is both unnecessary and impractical. As a retired [scientist]{.blue}, I urge you to abandon this reckless plan, for the sake of our [childrens health]{.red}. To do otherwise is unconscionable.</br></br>

<https://www.regulations.gov/comment/EPA-HQ-OA-2018-0259-19410>

## Example comment: Supporter {.smallish}

I sincerely hope that you can distance the EPA agency from [politically correct science]{.blue}. The only way to do this is to adopt the following [methodology]{.blue}.</br></br>

::: {.blue}
The scientific method demands that researchers' work be fully transparent.

Data must be made available for examination.

Theories must be open to challenge.

Disproven theories must be altered or set aside.

Barriers must be removed that would prevent valid conclusions from rising to the top.
:::
</br>

I for one am tired of hearing about how bad technologies are for the planet, [without any scientific proof]{.blue} that things today are different. Don't tell me that glarciers are melting and the ocean is rising a foot every 10 years. Remove yourselves from such statements and [give facts]{.blue} about such a change: volume of water, [how will it be measured, what could be causing it]{.blue}, [what can be done that is both environmentally and economically sound]{.red}.</br></br>

<https://www.regulations.gov/comment/EPA-HQ-OA-2018-0259-12305>

## Text mining methods

::: {.v-center}
- 21k comments is too many to analyze using traditional close reading
</br></br>

- 800 comments categorized as supporting/opposing by humans 
- remaining comments categorized using machine learning model
- natural language processing (NLP) to extract adjective-noun pairs
- keyword searches that account for synonyms and related concepts
:::

# Aims of science

## Rival views of the aims of science

@HicksScientificPracticesTheir2012; @HicksWhenVirtuesAre2022

- Two views of the aims of science
- Differ over relationship between epistemic and pragmatic aims

narrow view
  ~ Science has only epistemic aims; \
    practical applications are at best nice side effects, \
    must not override epistemic considerations
    </br></br>

broad view
  ~ Science has both epistemic and pragmatic aims; \
    relationship between the two is complex, context-dependent; \
    pragmatic considerations can take priority in some cases

## Rival views in ST public comments

![ \ ](../out/09_llr_top_filtered.png){}

## Rival views in ST public comments

::::: {.v-center}
:::: {.columns .center}
::: {.column}
[Opponents (liberals) use broad view]{.blue}
:::
::: {.column}
[Supporters (conservatives) use narrow view]{.red}
:::
::::

:::: {.columns}
::: {.column}
- Both epistemic & pragmatic aims</br></br>
- Public health as a primary aim of EPA/regulatory science
- ST would frustrate that aim
:::
::: {.column}
- Epistemic aims only; almost no discussion of practical concerns
- EPA/regulatory science as currently failing to achieve its epistemic aims
- ST would promote them
:::
::::
:::::

# Upshots

## Rival views as deep disagreement

::: {.v-center}
In the ST controversy, disagreement over the aims of science is **deep** in two senses:

axiomatic; self-evident
  ~ Each side treats their view as not in need of justification or defense

latent
  ~ Neither side seems to recognize the existence of disagreement

<br><br>

Just like [shared]{.blue} background assumptions can underpin consensus, \
[unshared]{.red} background assumptions can underpin disagreement
:::

## Cynicism and "Gold Standard Science"

:::: {.columns}
::: {.column}
For supporters of ST, science is

- secretive
- false
- phony
- deceptive
- shady
- bogus
- hidden
- fraudulent

:::
::: {.column}

[Cynical]{.red} frame for interpreting science

- Also shared by the mainstream science reform movement! 
- ST weaponized the rhetoric of this movement [@HicksOpenScienceReplication2023]
- Continued in the May 23, 2025, EO calling for a "return" to "Gold Standard Science" [@TollefsonTrumpsCallGold2025]

:::
::::

# {.placeholder-for-titleSlide visibility='uncounted'}

# References # {.unlisted visibility='uncounted' style="font-size:50%"}

### 2025-07-11 paper/paper.qmd
working on todos

@@
 
-abstract: "abstract"
+abstract: |
+    | *[update w/ conclusions]*
+    | 
+    | Environmental controversies are often framed in terms of "science vs.
+anti-science," with leftists and progressives typically claiming the side of science.
+However, their opponents often appeal to science as well.
+Hicks (2022) proposed that two rival views of the aims of science might be at play in these kinds of controversies.
+On the "narrow view" science has purely epistemic aims; while the "broad view" takes science to also have pragmatic or practical aims.
+    | 
+    | We argue that these rival views can be seen in the controversy over the proposed US EPA open data rule "Strengthening Transparency in Regulatory Science." The original justification for the rule appealed to the narrative of a science-wide replication crisis and the open science movement.
+In public comments, supporters of the rule likewise appealed to the narrow view of science, criticizing the titular "regulatory science" in epistemic rather than practical terms.
+While opponents of the rule appealed to the broad view, connecting science to the practical aim of protecting human health.
+    | 
+    | We analyze approximately 21,000 public comments on the proposed rule, combining manual coding with machine learning to classify every comment as supporting or opposing, and using natural language processing (NLP) and text mining methods to compare how supporters and opponents talk about science in their comments.
 
@@
 
-Something like Hicks' two views appears in philosophical debates over the argument from inductive risk. *[acceptance vs. belief]*
+Something like Hicks' two views of the aims of science appears in philosophical debates over the argument from inductive risk.
+First, the argument itself treats the non-epistemic consequences of research as on a par with the epistemic consequences, similarly to the way the broad view does not prioritize epistemic aims over pragmatic ones.
+Next, in his objection to Rudner's original formulation of the argument, @LeviMustScientistMake1960 claimed that Rudner (and other proponents of the argument) relied on an assumption that "To choose to accept a hypothesis as true (or to believe that H is true) is equivalent to choosing to act on the basis of H relative to some specific objective P" (348), which he associated with decision theory and, later, behaviorism [@LeviSeriousnessMistakes1962].
+Denying this assumption, @LeviMustScientistMake1960 distinguished the "theoretical" or epistemic objective of "accepting true answers to questions as true" from "practical" objectives (350).
+Similarly, @LeviSeriousnessMistakes1962 defends the view that the aim of scientific inquiry is to "replace doubt by true belief" (48).
+This aim might include either "seek[ing] the truth and nothing but the truth" or, more broadly, "certain desirable properties in addition to truth" such as "simplicity, explanatory power," and presumably other epistemic virtues (48).
+But, on either construal, "replac[ing] doubt by true belief" is "theoretical" or epistemic and "non-behavioral" (48).
+In other words, Levi objected to the argument from inductive risk by articulating the narrow view.
+*[Franco cites]*.
+Finally, in response to similar critiques of *[Douglas']* version of the argument *[eg, Lacey; Mayo?]*, critics of the value-free ideal adopted *[Cohen's?]* distinction between belief (in a claim as true) and acceptance (for practical purposes), explicitly scoping their arguments to acceptance *[cites]*.
+This move in the debate concedes Levi's rejection of the "behaviorist" identification of belief and acceptance — but also rejects the narrow view's purely epistemic aims for inquiry, placing defenders and critics of VFI on opposite sides of the two views of the aims of science.
 
@@
 *[secret science act]*
+
+*[Gold Standard Science EO]*
 
@@
 
-*[set blue L to like 40]*
-
 ![15 most common adjective-noun bigrams, by support for ST (rows of panels) and codings (columns of panels).
@@
 
-*[already using red/blue for health/science]*
-
 ![15 most distinctive terms (unigrams) for supporters and opponents, by coding.
@@
 
-We argue that the results above can be fruitfully understood in terms of @HicksWhenVirtuesAre2022 account of two views of the aims of science.
+We argue that the results above can be fruitfully understood in terms of Hicks' -@HicksWhenVirtuesAre2022 account of two views of the aims of science.
 Specifically, we claim that supporters deployed the narrow view, while opponents deployed the broad view.
@@
 We therefore take the aims approach to be a better fit for analyzing this particular controversy.
+
+Yet another alternative account of this controversy would be in terms of the value-free ideal, understood as the view that non-epistemic values must not influence the "internal" or "core" stages of scientific inquiry — analyzing evidence, evaluating hypotheses — but may permissibly influence "earlier" and "later" stages of inquiry — such as posing research questions and utilizing research findings in policymaking.
+On this reading, supporters of ST perceive EPA/regulatory science as frequently violating the value-free ideal — perhaps environmentalist or "leftist" values are influencing the evaluation of hypotheses — while opponents do not.^[Thanks to [redacted]<!-- Soazig Le Bihan --> for suggesting a version of this objection.
+[Their] version did not tie the internal/external influence distinction to the value-free ideal.]
+
+This account is not as much of an alternative to our aims-based account as it might seem.
+@HicksWhenVirtuesAre2022 argues that the narrow view is often assumed by defenders of the value-free ideal (VFI), and so for supporters of ST there's not much difference between "failing to achieve narrowly epistemic aims" and "violating VFI." However, for opponents of ST, this alternative account would require parsing their comments on ST into VFI's three-stage model of inquiry, and showing that they view non-epistemic values as influencing only the "external" stages (research questions and downstream applications).
+Concerns that ST would enable "political interference" — presidential appointees influencing the evaluation of hypotheses and conclusions reached by scientific reports — can be read as concerns about VFI violations.
+But other concerns would be much more difficult to parse in terms of the three-stage model assumed by VFI.
+
+The VFI-based account of the ST controversy also does not fit well with empirical research on the relationship among VFI, cynicism about science, and trust in science among members of the general public in the US.
+@HicksDevelopingMeasuresPublic2024 find that members of the public generally agree with VFI, and the average score along a composite measure of "objectivity" is relatively high.
+But they also find that this measure does not predict trust in science.
+Instead a different composite measure, "cynicism," predicts trust in science,^[Specifically, a high score on the cynicism scale is a good predictor of *lower* trust in science.] and cynicism and objectivity are not correlated with each other.
+@HicksDevelopingMeasuresPublic2024 also connect the cynicism measure to the failure of science to achieve its aims.
+Below we discuss how supporters of ST often express significant cynicism towards EPA/regulatory science.
 

### 2025-07-11 scripts/11_keyword_occurrence.R

@@ -49,11 +49,11 @@ map(coded, count_tokens) |>
     scale_x_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
     # guides(x = guide_axis_logticks()) +
     annotation_logticks(sides = 'b') +
-    scale_color_brewer(palette = 'Set1') +
     facet_wrap(vars(coding), 
                ncol = 1, 
                scales = 'free_y', 
-               axes = 'all_x')
+               axes = 'all_x') +
+    scale_color_brewer(palette = 'Dark2')
 
 ggsave(here(out_dir, '11_len.png'), 
        height = 3, width = 4, bg = 'white', scale = 2)
@@ -168,7 +168,8 @@ llr_df |>
     geom_point(aes(y = llr)) +
     scale_x_reordered() +
     coord_flip() +
-    facet_wrap(vars(coding), ncol = 3, scales = 'free_y')
+    facet_wrap(vars(coding), ncol = 3, scales = 'free_y') +
+    scale_color_brewer(palette = 'Dark2')
 
 ggsave(here(out_dir, '11_llr.png'), 
        height = 3, width = 4, bg = 'white', scale = 2)
